---
title: "Generated code"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generated code}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

```{r setup, include = FALSE}
library(ckbplotr)
```

The `shape_plot()` and `forest_plot()` functions return both a plot and the code used to generate plot (which will be shown in the RStudio Viewer pane). Examples of generated code are shown below.

## Shape plot

```{r, shape-example}
my_results <- data.frame(
  x   = c(  12,   14, 15.5,   18),
  est = c(0.05, 0.21, 0.15, 0.32),
  se  = c(0.05, 0.05, 0.05, 0.05)
)

shape <- shape_plot(my_results,
                    xlims        = c(10, 20),
                    ylims        = c(0.75, 2),
                    exponentiate = TRUE,
                    printplot    = FALSE)

```

The code now stored in `shape$code` and shown in the RStudio Viewer pane is:
```{r, shape-example-code, results ='asis', echo = FALSE}
cat("",
    "```",
    shape$code,
    "```",
    sep = "\n")
```


## Forest plot

```{r, forest-example}
my_results <- data.frame(
  subgroup   = c("men", "women", "35_49", "50_64", "65_79"),
  est        = c( 0.45,    0.58,    0.09,    0.35,     0.6),
  se         = c( 0.07,    0.06,    0.06,    0.05,    0.08)
)

forest <- forest_plot(my_results, printplot = FALSE)
```

The code now stored in `forest$code` and shown in the RStudio Viewer pane is:
```{r, forest-example-code, results ='asis', echo = FALSE}
cat("",
    "```",
    forest$code,
    "```",
    sep = "\n")
```
