---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.width = 8,
  fig.height = 7,
  dpi = 100,
  out.width = "100%"
)
```
# ckbplotr

`ckbplotr` provides functions to help create and style plots in R. It is being
developed by, and primarily for, [China Kadoorie Biobank](http://www.ckbiobank.org) 
researchers.

## Installation

### Directly from github
The latest development version of `ckbplotr` can be installed directly from 
github using devtools.
```{r install from github, eval = FALSE}
install.packages('devtools')
devtools::install_github('neilstats/ckbplotr')
```

### Or from source package
`ckbplotr` can also be installed from its source package. The R packages 
`ggplot2`, `magrittr`, `readr`, `tibble`, `dplyr`, `purrr` and `rlang` must 
first be installed. These are part of the collection of tidyverse packages.
```{r install tidyverse, eval = FALSE}
# The easiest way is to install the whole tidyverse:
install.packages("tidyverse")

# # Or install just these packages:
# install.packages(c("ggplot2", "readr", "dplyr", "purrr"))
```
Then `ckbplotr` can be installed from its source package using the code:
```{r install from source, eval = FALSE}
install.packages("ckbplotr.tar.gz", repos = NULL, type = "source")
```
Or, in RStudio, open the "Tools" menu and select "Install Packages...".
In the "Install from..." box select "Package Archive File", and in the
"Package archive" box browse to the ckbplotr.tar.gz file.

The source package for the latest release version is available [here](https://github.com/neilstats/ckbplotr/releases/latest).

## The make_shape_plot function

The `make_shape_plot` function creates a plot of estimates and CIs against risk
factor levels using the `ggplot2` package.

The function returns a named list containing:

* plot: the plot
* code: ggplot2 code to generate the plot

The function uses the `plot_like_ckb` function to modify the theme and control
the look of the plot.

In RStudio, the ggplot2 code used to generate the plot will be shown in the 'Viewer'
pane. This code can be copied and further edited to adapt the plot.


### Examples
Given a data frame of estimates and standard errors (to be plotted on the y axis),
and risk factor levels (to be plotted on the x axis), a plot can be created.
```{r make_shape_plot-example-1}
library(ggplot2)
library(ckbplotr)

results <- data.frame(
  est = c(1, 1.25, 1.5, 1.75, 1.1, 1.35, 1.75, 2),
  se = c(0.08, 0.02, 0.07, 0.15, 0.05, 0.02, 0.07, 0.1),
  rf = c(20, 28, 35, 50, 18.5, 25, 32, 47),
  n = c(109, 103, 143, 104, 140, 134, 127, 99),
  is_female = c(0, 0, 0, 0, 1, 1, 1, 1)
)

plot <- make_shape_plot(results[results$is_female == 1,],
                        col.x        = "rf",
                        col.estimate = "est",
                        col.stderr   = "se",
                        col.n        = "n",
                        xlims        = c(15,50),
                        ylims        = c(0.5, 3),
                        scalepoints  = TRUE,
                        title        = NULL)
```

The `col.group` argument can be supplied to plot results for different groups (in
shades of grey).
```{r make_shape_plot-example-2}
plot <- make_shape_plot(results,
                        col.x        = "rf",
                        col.estimate = "est",
                        col.stderr   = "se",
                        col.n        = "n",
                        col.group    = "is_female",
                        xlims        = c(15,50),
                        ylims        = c(0.5, 3),
                        scalepoints  = TRUE,
                        title        = NULL)
```




## The plot_like_ckb function

The `plot_like_ckb` function does three things to a ggplot2 plot:

1. modifies the theme (i.e. the overall appearance)
2. extends the plotting area, hides the default axis lines and manually adds 
axis lines (so that you can have a custom sized gap between the plotting area and the axes)
3. applies a fixed aspect ratio

### Examples

Make a scatter plot with `ggplot`.
```{r a-plot}
library(ggplot2)
library(ckbplotr)

plot <- ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(size = 1)
plot
```

Then apply the `plot_like_ckb` function to the plot.
```{r plot_like_ckb-example-1}
plot_like_ckb(plot = plot, xlims = c(0, 8), ylims = c(10, 50))
```

If you do not want a gap between the axes, set `gap = c(0, 0)`.
```{r plot_like_ckb-example-2}
plot_like_ckb(plot = plot, xlims = c(0, 8), ylims = c(10, 50), gap = c(0, 0))
```

If you wish to override some aspect of the theme applied by `plot_like_ckb`, 
then this can be done by adding a theme after the function.
```{r plot_like_ckb-example-3}
plot_like_ckb(plot = plot, xlims = c(0, 8), ylims = c(10, 50)) +
  theme(axis.title = element_text(colour = "red", face = "plain"))
```

If any data points you are plotting fall outside the axes, then they will still
be drawn and may show up in places such as the axes, the legend, the plot title, 
or the plot margins. There is also no warning if data points fall outside the whole plot area.

So it is best to check that your `xlim` and `ylim` values are suitable for your 
data before using the function.
```{r plot_like_ckb-example-4}
# The xlim and ylim ranges are too narrow
plot_like_ckb(plot = plot, xlims = c(0, 4), ylims = c(20, 50))
```


A bar chart.
```{r bar-chart-example}
barplot <- ggplot(mpg, aes(class)) + geom_bar()
plot_like_ckb(barplot, xlims = c(0.5, 7.5), ylims = c(0, 70), gap = c(0.025, 0.005), ratio = 0.5) +
  theme(axis.ticks.x = element_blank())
```



## The make_forest_plot function
The `make_forest_plot` function creates a forest plot using the `ggplot2` graphics 
package. It is available as part of the `ckbplotr` package.

The function returns a named list containg:

* plot: the plot
* data: a data frame from which the plot is generated
* code: ggplot2 code to generate the plot

In RStudio the ggplot2 code used to generate the plot will be shown in the 
'Viewer' pane. If modifications are needed to the plot, then this code can be
copied, edited, and run as needed.


### Expected warning
You may get the following warning when using this function.

> Vectorized input to `element_text()` is not officially supported.
> Results may be unexpected or may change in future versions of ggplot2. 

This is related to how some headings used to label the rows of the forest plot are
made bold. This should not be a problem if the forest plot produced appears as expected.
(This hack is discussed [here](https://github.com/tidyverse/ggplot2/issues/3492).)


### The headings data frame
The data frame specified in the `headings` argument allows the use of headings 
to label the rows of the forest plot.

This must data frame contain columns 'heading1', 'heading2' and 'heading3',
which can be set to NA if not needed for a particular row. For example, if you 
do not need a third level of heading for a particular row, then set `heading3=NA`
for that row. If you only need one level of headings, then set `heading2=NA` and
`heading3=NA` for every row.

The data set must also contain a column with the same name as a column in each 
of the results data sets. This column is used to match headings to the correct 
results, and is specified in the argument `col.key`.

The argument `rows` specifies which results should be included in the plot, by
giving some headings from the heading1 column of the headings data frame.

(The order given in the `rows` argument decides the order in which the heading1 
labels will be in the plot (top to bottom). The heading2 and heading3 labels
will be in the order that they are in the headings data frame.)


### Examples

First of all, do your analyses and put the results into data frames 
(one data frame for each forest plot column). Here we generate some example results.
```{r example forest plot results}
set.seed(57911624)
exampleresults <- function(){
  data.frame(variable = c('nmr_l_vldl_p', 'nmr_m_vldl_p', 'nmr_s_vldl_p',
                          'nmr_idl_p',
                          'nmr_l_ldl_p', 'nmr_m_ldl_p', 'nmr_s_ldl_p',
                          'nmr_l_vldl_tg', 'nmr_m_vldl_tg', 'nmr_s_vldl_tg',
                          'nmr_idl_tg',
                          'nmr_l_ldl_tg', 'nmr_m_ldl_tg', 'nmr_s_ldl_tg',
                          'nmr_l_vldl_c', 'nmr_m_vldl_c', 'nmr_s_vldl_c',
                          'nmr_idl_c',
                          'nmr_l_ldl_c', 'nmr_m_ldl_c', 'nmr_s_ldl_c'),
             estimate = rnorm(21, 0, 0.02),
             stderr   = 0.012 + abs(rnorm(21, 0, 0.015)),
             n        = round(runif(21, 100, 2000)),
             nb       = round(runif(21, 100, 2000)))
}
resultsA <- exampleresults()
resultsB <- exampleresults()
resultsC <- exampleresults()
resultsD <- exampleresults()
resultsE <- exampleresults()
```

```{r make_forest_plot-example-1}
library(ckbplotr)

forestplot <- make_forest_plot(cols         = list(resultsA, resultsB),
                               exponentiate = TRUE,
                               colnames     = c("Analysis A", "Analysis B"),
                               col.key      = "variable")
```

To use headings, create a data frame of headings and specify arguments `headings` and
`rows`.
```{r forest plot headings, eval=TRUE, echo=FALSE}
# Example headings for every row
# This could be loaded from a csv containing headings that you use often
headings <- data.frame(heading1 = rep(c("Lipoprotein particle concentration",
                                        "Triglycerides concentration",
                                        "Cholesterol concentration"), each = 7),
                       heading2 = rep(c("VLDL", "VLDL", "VLDL",
                                        "IDL",
                                        "LDL", "LDL", "LDL"), times = 3),
                       heading3 = rep(c("Large", "Medium", "Small",
                                        NA,
                                        "Large", "Medium", "Small") , times = 3),
                       variable = c('nmr_l_vldl_p', 'nmr_m_vldl_p', 'nmr_s_vldl_p',
                                    'nmr_idl_p',
                                    'nmr_l_ldl_p', 'nmr_m_ldl_p', 'nmr_s_ldl_p',
                                    'nmr_l_vldl_tg', 'nmr_m_vldl_tg', 'nmr_s_vldl_tg',
                                    'nmr_idl_tg',
                                    'nmr_l_ldl_tg', 'nmr_m_ldl_tg', 'nmr_s_ldl_tg',
                                    'nmr_l_vldl_c', 'nmr_m_vldl_c', 'nmr_s_vldl_c',
                                    'nmr_idl_c',
                                    'nmr_l_ldl_c', 'nmr_m_ldl_c', 'nmr_s_ldl_c'))
headings
```

```{r make_forest_plot-example-2}
forestplot <- make_forest_plot(headings         = headings,
                               rows             = c("Lipoprotein particle concentration",
                                                    "Triglycerides concentration"),
                               cols             = list(resultsA, resultsB),
                               exponentiate     = TRUE,
                               colnames         = c("Analysis A", "Analysis B"),
                               col.key          = "variable",
                               ci.delim         = " - ",
                               xlim             = c(0.9, 1.1),
                               xticks           = c(0.9, 1, 1.1),
                               blankrows        = c(1, 1, 0, 1),
                               scalepoints      = TRUE,
                               pointsize        = 3,
                               col.left         = c("n"),
                               col.left.space   = c(0.02),
                               col.left.heading = c("No. of\nevents"),
                               col.right.space  = 0.02,
                               col.heading.space = 1.5,
                               heading.space    = 2,
                               plot.space       = 8)
```


## make_jasper_forest_plot function
The `make_jasper_forest_plot` function requires the in-house Jasper package, 
which is not publicly available.

