---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# ckbplotr

`ckbplotr` provides functions to help create and style plots in R. It is being
developed by, and primarily for, [China Kadoorie Biobank](http://www.ckbiobank.org) 
researchers.

## Installation
The latest development version of `ckbplotr` can be installed directly from 
github using devtools.
```{r, eval = FALSE}
install.packages('devtools')
devtools::install_github('neilstats/ckbplotr')
```

`ckbplotr` can also be installed from its source package. The R packages 
`ggplot2`, `magrittr`, `readr`, `tibble`, `dplyr`, `purrr` and `rlang` must 
first be installed. These are part of the collection of tidyverse packages.
```{r, eval = FALSE}
# The easiest way is to install the whole tidyverse:
install.package("tidyverse")

# # Or install just these packages:
# install.packages(c("ggplot2", "readr", "dplyr", "purrr"))
```
Then `ckbplotr` can be installed from its source package using the code:
```{r, eval = FALSE}
install.packages("ckbplotr.tar.gz", repos = NULL, type = "source")
```
Or, in RStudio, open the "Tools" menu and select "Install Packages...".
In the "Install from..." box select "Package Archive File", and in the
"Package archive" box browse to the ckbplotr.tar.gz file.




## The make_shape_plot function

The `make_shape_plot` function creates a plot of estimates and CIs against risk
factor levels using the `ggplot2` package.

The function returns a named list containing:

* plot: the plot
* code: ggplot2 code to generate the plot

The function uses the `plot_like_ckb` function to modify the theme and control
the look of the plot.

In RStudio, the ggplot2 code used to generate the plot will be shown in the 'Viewer'
pane. This code can be copied and further edited to adapt the plot.


### Examples
Given a data frame of estimates and standard errors (to be plotted on the y axis),
and risk factor levels (to be plotted on the x axis), a plot can be created.
```{r}
library(ggplot2)
library(ckbplotr)

results <- data.frame(
  est = c(1, 1.25, 1.5, 1.75, 1.1, 1.35, 1.75, 2),
  se = c(0.08, 0.02, 0.07, 0.15, 0.05, 0.02, 0.07, 0.1),
  rf = c(20, 28, 35, 50, 18.5, 25, 32, 47),
  n = c(109, 103, 143, 104, 140, 134, 127, 99),
  is_female = c(0, 0, 0, 0, 1, 1, 1, 1)
)

plot <- make_shape_plot(results[results$is_female == 1,],
                        col.x        = "rf",
                        col.estimate = "est",
                        col.stderr   = "se",
                        col.n        = "n",
                        xlims        = c(15,50),
                        ylims        = c(0.5, 3),
                        scalepoints  = TRUE,
                        title        = NULL)
```

The `col.group` argument can be supplied to plot results for different groups (in
shades of grey).
```{r}
plot <- make_shape_plot(results,
                        col.x        = "rf",
                        col.estimate = "est",
                        col.stderr   = "se",
                        col.n        = "n",
                        col.group    = "is_female",
                        xlims        = c(15,50),
                        ylims        = c(0.5, 3),
                        scalepoints  = TRUE,
                        title        = NULL)
```




## The plot_like_ckb function

The `plot_like_ckb` function does three things to a ggplot2 plot:

1. modifies the theme (i.e. the overall appearance)
2. extends the plotting area, hides the default axis lines and manually adds 
axis lines (so that you can have a custom sized gap between the plotting area and the axes)
3. applies a fixed aspect ratio

### Examples

Make a scatter plot with `ggplot`.
```{r}
library(ggplot2)
library(ckbplotr)

plot <- ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(size = 1)
plot
```

Then apply the `plot_like_ckb` function to the plot.
```{r}
plot_like_ckb(plot = plot, xlims = c(0, 8), ylims = c(10, 50))
```

If you do not want a gap between the axes, set `gap = c(0, 0)`.
```{r}
plot_like_ckb(plot = plot, xlims = c(0, 8), ylims = c(10, 50), gap = c(0, 0))
```

If you wish to override some aspect of the theme applied by `plot_like_ckb`, 
then this can be done by adding a theme after the function.
```{r}
plot_like_ckb(plot = plot, xlims = c(0, 8), ylims = c(10, 50)) +
  theme(axis.title = element_text(colour = "red", face = "plain"))
```

If any data points you are plotting fall outside the axes, then they will still
be drawn and may show up in places such as the axes, the legend, the plot title, 
or the plot margins. There is also no warning if data points fall outside the whole plot area.

So it is best to check that your `xlim` and `ylim` values are suitable for your 
data before using the function.
```{r}
# The xlim and ylim ranges are too narrow
plot_like_ckb(plot = plot, xlims = c(0, 4), ylims = c(20, 50))
```


A bar chart.
```{r}
barplot <- ggplot(mpg, aes(class)) + geom_bar()
plot_like_ckb(barplot, xlims = c(0.5, 7.5), ylims = c(0, 70), gap = c(0.025, 0.005), ratio = 0.5) +
  theme(axis.ticks.x = element_blank())
```



## make_jasper_forest_plot function
The `make_jasper_forest_plot` function requires the in-house Jasper package, 
which is not publicly available.

