---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.width = 4,
  fig.height = 4,
  fig.align = "center",
  dpi = 300)
library(ckbplotr)
library(grid)
library(gridExtra)
```
# ckbplotr
<!-- badges: start -->
[![R build status](https://github.com/neilstats/ckbplotr/workflows/R-CMD-check/badge.svg)](https://github.com/neilstats/ckbplotr/actions)
[![Codecov test coverage](https://codecov.io/gh/neilstats/ckbplotr/branch/master/graph/badge.svg)](https://codecov.io/gh/neilstats/ckbplotr?branch=master)
<!-- badges: end -->

`ckbplotr` provides functions to help create and style plots in R. It is being
developed by, and primarily for, [China Kadoorie Biobank](http://www.ckbiobank.org) 
researchers.

## Installation

### Directly from github
The latest version of `ckbplotr` can be installed directly from 
github using the `remotes` package.
```{r install from github, eval = FALSE}
install.packages('remotes')
remotes::install_github('neilstats/ckbplotr')
```

If you get an error that reads "Error: Failed to install 'unknown package' from GitHub: HTTP error 404. No commit found for the ref master" then make sure to update to the latest version of the `remotes` package, or try `remotes::install_github('neilstats/ckbplotr@main')`.

If you get an error that reads "Error: (converted from warning) package 'ggplot2' was built under R version ..." you can avoid this by first running `Sys.setenv("R_REMOTES_NO_ERRORS_FROM_WARNINGS" = "true")`. (This is a [known issue](https://github.com/r-lib/remotes/issues/403) with the `remotes` package.) 

### Or from source package
`ckbplotr` can also be installed from its source package. The R packages  `ggplot2`, `magrittr`, `readr`, `tibble`, `dplyr`, `purrr`,  `rlang`, and `ggtext` must  first be installed.
```{r install tidyverse, eval = FALSE}
# The easiest way is to install the whole tidyverse and ggtext:
install.packages("tidyverse", "ggtext")

# # Or install just these packages:
# install.packages(c("ggplot2", "readr", "dplyr", "purrr", "ggtext"))
```
Then `ckbplotr` can be installed from its source package using the code:
```{r install from source, eval = FALSE}
install.packages("ckbplotr.tar.gz", repos = NULL, type = "source")
```
Or, in RStudio, open the "Tools" menu and select "Install Packages...".
In the "Install from..." box select "Package Archive File", and in the
"Package archive" box browse to the ckbplotr.tar.gz file.

The source package for the latest release version is available [here](https://github.com/neilstats/ckbplotr/releases/latest).




## Shape plots
`make_shape_plot()` creates a plot of estimates and CIs against risk factor levels using the [ggplot2](https://ggplot2.tidyverse.org/) graphics package. The function returns both a plot and the ggplot2 code used to create the plot. In RStudio the ggplot2 code used to create the plot will be shown in the Viewer pane (with syntax highlighting if the [highlights](https://cran.r-project.org/package=highlight) package is installed).

```{r make_shape_plot-example-1, echo = FALSE, out.width = "50%"}
results <- data.frame(
  est = c(1, 1.25, 1.5, 1.75, 1.1, 1.35, 1.75, 2),
  se = c(0.08, 0.02, 0.07, 0.15, 0.05, 0.02, 0.07, 0.1),
  rf = c(20, 28, 35, 50, 18.5, 25, 32, 47),
  n = c(109, 103, 143, 104, 140, 134, 127, 99),
  is_female = c(0, 0, 0, 0, 1, 1, 1, 1)
) %>% 
  dplyr::filter(is_female == 0)

plot <- make_shape_plot(results,
                        col.x        = "rf",
                        col.estimate = "est",
                        col.stderr   = "se",
                        col.n        = "n",
                        col.group    = "is_female",
                        xlims        = c(15,50),
                        ylims        = c(0.5, 2.5),
                        scalepoints  = TRUE,
                        title        = NULL,
                        ciunder      = TRUE,
                        envir        = environment())
```





## Forest plots
`make_forest_plot()` creates a forest plot using the [ggplot2](https://ggplot2.tidyverse.org/) graphics package. The function returns both a plot and the ggplot2 code used to create the plot. In RStudio the code used to create the plot will be shown in the Viewer pane (with syntax highlighting if the [highlights](https://cran.r-project.org/package=highlight) package is installed).

```{r example-forest-plot, echo = FALSE, warning = FALSE, fig.width = 8, fig.height = 6, out.width = "90%"}
set.seed(57911624)
exampleresults <- function(){
  data.frame(variable = c('nmr_l_vldl_p', 'nmr_m_vldl_p', 'nmr_s_vldl_p',
                          'nmr_idl_p',
                          'nmr_l_ldl_p', 'nmr_m_ldl_p', 'nmr_s_ldl_p',
                          'nmr_l_vldl_tg', 'nmr_m_vldl_tg', 'nmr_s_vldl_tg',
                          'nmr_idl_tg',
                          'nmr_l_ldl_tg', 'nmr_m_ldl_tg', 'nmr_s_ldl_tg',
                          'nmr_l_vldl_c', 'nmr_m_vldl_c', 'nmr_s_vldl_c',
                          'nmr_idl_c',
                          'nmr_l_ldl_c', 'nmr_m_ldl_c', 'nmr_s_ldl_c'),
             estimate = rnorm(21, 0, 0.02),
             stderr   = 0.012 + abs(rnorm(21, 0, 0.015)),
             n        = round(runif(21, 100, 2000)),
             nb       = round(runif(21, 100, 2000)))
}
resultsA <- exampleresults()
resultsB <- exampleresults()
resultsC <- exampleresults()
resultsD <- exampleresults()
resultsE <- exampleresults()

mylabels <- data.frame(heading1 = rep(c("Lipoprotein particle concentration",
                                        "Triglycerides concentration",
                                        "Cholesterol concentration"), each = 7),
                       heading2 = rep(c("VLDL", "VLDL", "VLDL",
                                        "IDL",
                                        "LDL", "LDL", "LDL"), times = 3),
                       heading3 = rep(c("Large", "Medium", "Small",
                                        NA,
                                        "Large", "Medium", "Small") , times = 3),
                       variable = c('nmr_l_vldl_p', 'nmr_m_vldl_p', 'nmr_s_vldl_p',
                                    'nmr_idl_p',
                                    'nmr_l_ldl_p', 'nmr_m_ldl_p', 'nmr_s_ldl_p',
                                    'nmr_l_vldl_tg', 'nmr_m_vldl_tg', 'nmr_s_vldl_tg',
                                    'nmr_idl_tg',
                                    'nmr_l_ldl_tg', 'nmr_m_ldl_tg', 'nmr_s_ldl_tg',
                                    'nmr_l_vldl_c', 'nmr_m_vldl_c', 'nmr_s_vldl_c',
                                    'nmr_idl_c',
                                    'nmr_l_ldl_c', 'nmr_m_ldl_c', 'nmr_s_ldl_c'))

forestplot <- make_forest_plot(row.labels        = mylabels,
                               rows              = c("Lipoprotein particle concentration",
                                                     "Triglycerides concentration"),
                               panels            = list(resultsA, resultsB),
                               exponentiate      = TRUE,
                               panel.names       = c("Analysis A", "Analysis B"),
                               col.key           = "variable",
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 1, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 3,
                               col.left          = c("n"),
                               col.left.space    = c(0.02),
                               col.left.heading  = c("No. of\nevents"),
                               col.right.space   = 0.02,
                               col.heading.space = 1.5,
                               envir             = environment())
```

#### make_jasper_forest_plot function
`make_jasper_forest_plot()` requires the in-house Jasper package,  which is not publicly available.




## Other plots

`plot_like_ckb()` does three things to a [ggplot2](https://ggplot2.tidyverse.org/) plot:

1. applies a CKB theme (i.e. change the overall appearance)
2. extends the plotting area and manually adds axis lines (so that you can have a custom sized gap between the plotting area and the axes)
3. applies a fixed aspect ratio

```{r a-plot, echo = FALSE, fig.width=8, out.width = "80%"}
plot <- ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(size = 1)
ckbplot <- plot_like_ckb(plot = plot, xlims = c(0, 8), ylims = c(10, 50))
grid.arrange(plot, ckbplot, ncol = 2)
```

