---
title: "CKB style with ggplot"
output:
  rmarkdown::html_vignette:
    fig_width: 4
    fig_height: 4
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{CKB style with ggplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 4,
  fig.height = 4,
  fig.align = "center",
  out.width = '50%'
)
```

```{r setup, include = FALSE}
library(ckbplotr)
```

`+ ckb_style()` does three things to a ggplot2 plot:

1. applies a CKB theme (i.e. change the overall appearance)
2. extends the plotting area and manually adds axis lines (so that you can have a custom sized gap between the plotting area and the axes)
3. applies a fixed aspect ratio

## Examples with a scatter plot
Make a scatter plot with `ggplot`.
```{r a-plot}
plot <- ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(size = 1)
plot
```

Use `+ ckb_style()` to apply a CKB theme, add separated axis lines and fix the aspect ratio.
```{r plot_like_ckb-example-0}
plot + ckb_style()
```

Or apply just a CKB theme.
```{r theme}
plot + theme_ckb()
```


### Axis limits
Set axis limits.
```{r plot_like_ckb-example-1}
plot + ckb_style(xlims = c(0, 8), ylims = c(10, 50))
```

If you do not want a gap between the axes, set `gap = c(0, 0)`.
```{r plot_like_ckb-example-2}
plot + ckb_style(xlims = c(0, 8), ylims = c(10, 50), gap = c(0, 0))
```

### Aspect ratio and panel sizes
Change the aspect ratio of the plot.
```{r aspect-ratio, fig.width = 6}
plot + ckb_style(xlims = c(0, 8),
                 ylims = c(10, 50),
                 ratio = 0.3)
```

Set the width of the plot (`width` controls the length of the x axis).
```{r width, fig.width = 6}
plot + ckb_style(xlims = c(0, 8),
                 ylims = c(10, 50),
                 ratio = 1.5,
                 width = unit(4, "cm"))
```



### Modifying the appearance of the plot
The `colour` arguments of `theme_ckb()` and `ckb_style()` can be used to change the colour of the non-data components of the plot.
```{r colour}
plot + ckb_style(xlims = c(0, 8),
                 ylims = c(10, 50),
                 colour = "darkred")
```

The `plot.margin` arguments of `theme_ckb()` and `ckb_style()` can be used to adjust the margin around the plot.
```{r plotmargin}
plot + ckb_style(xlims = c(0, 8),
                 ylims = c(10, 50),
                 plot.margin = margin(2, 2, 2, 2, unit = "cm"))
```

If you wish to override some aspect of the theme applied by `ckb_style()` or `theme_ckb()`, 
then this can be done by adding a theme after `+ ckb_style()`,
```{r plot_like_ckb-example-3}
plot + ckb_style(xlims = c(0, 8),
                 ylims = c(10, 50)) +
  theme(axis.title = element_text(colour = "red", face = "plain"))
```


### Data points outside axis limits
There is no warning if data points fall outside the axis limits. So check that your `xlim` and `ylim` values are suitable for your data before using the function.
```{r plot_like_ckb-example-4}
# The xlim and ylim ranges are too narrow
plot + ckb_style(xlims = c(0, 4), ylims = c(20, 50))
```


## Bar chart example
```{r bar-chart-example, fig.width = 6, out.width = '70%'}
ggplot(mpg, aes(class)) + 
  geom_bar() +
  ckb_style(xlims = c(0.5, 7.5),
            ylims = c(0, 70),
            gap = c(0.025, 0.005),
            ratio = 0.5) +
  theme(axis.ticks.x = element_blank())
```
