---
title: "Get started with ckbplotr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with ckbplotr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 7,
  fig.align = "center"
)
```

```{r setup, include = F}
library(ckbplotr)
```

## Introduction
The `ckbplotr` R package provides functions to help create and style plots in R using the [ggplot2](https://ggplot2.tidyverse.org/) graphics package.

## CKB style
Use `+ ckb_style()` to apply CKB style to a ggplot. See [CKB style ggplot](ckb_style.html) for details.

## Plots
Create plots of estimates against risk factor levels using `shape_plot()`. See [Shape plots](shape_plots.html) for details.

Create forest plots using `forest_plot()`. See [Forest plots](forest_plots.html) for details.

## Saving
See [Save plots to files](save_plots.html).

## More

Additional examples and tutorials for using the package can be found at <https://neilstats.github.io/ckbplotr-book>.
