---
title: "Get started with ckbplotr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with ckbplotr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 7,
  fig.align = "center"
)
```

```{r setup, include = F}
library(ckbplotr)
```
## Shape plots
`shape_plot()` creates a plot of estimates and CIs against risk factor levels using the `ggplot2` package. The function returns both a plot and the ggplot2 code used to create the plot. In RStudio the ggplot2 code used to create the plot will be shown in the  Viewer pane.

The function uses `ckb_style()` to modify the theme and control the look of the plot.

### Example data
```{r}
ckbplotr_shape_data
```

### Simple example
Given a data frame of estimates and standard errors (to be plotted on the y axis),
and risk factor levels (to be plotted on the x axis), a plot can be created.
```{r simple-example, fig.height = 4, fig.width = 4, out.width = "50%"}
shape_plot(ckbplotr_shape_data[ckbplotr_shape_data$is_female == 0,],
           col.x        = "rf",
           col.estimate = "est",
           col.stderr   = "se",
           col.n        = "n",
           xlims        = c(15, 50),
           ylims        = c(0.5, 3),
           scalepoints  = TRUE)
```

For more details on customising these plots, see ["More with shape plots"](shape_plot.html).

*Narrow confidence interval lines may become hidden in plots. Please check your final output carefully and see ["Confidence interval lines in shape plots"](shape_confidence_intervals.html) for methods for handle this.*

## Forest plots
`forest_plot()` creates a forest plot using the [ggplot2](https://ggplot2.tidyverse.org/) graphics package. The function returns both a plot and the ggplot2 code used to create the plot. In RStudio the code used to create the plot will be shown in the Viewer pane.

*Narrow confidence interval lines may become hidden in forest plots. Please check your final output carefully and see ["Confidence interval lines in forest plots"](forest_confidence_intervals.html) for methods for handle this.*

### Prepare data
First of all, do your analyses and put the results into data frames (one data frame for each forest plot panel). Here we take subsets of the example data provided in the package.
```{r example forest plot results}
resultsA <- dplyr::filter(ckbplotr_forest_data, name == "A")
resultsB <- dplyr::filter(ckbplotr_forest_data, name == "B")

head(resultsA[,c("variable", "estimate", "stderr", "n")])
```


### Simple forest plot
The `panels` argument is a list of data frames with results to be plotted, the `col.key` argument is the name of the column that defines which results should be plotted on the same row, and the `exponentiate` argument is a logical value telling the function we are providing estimates on the log scale.

By default, the function assumes that estimates and standard errors are in columns named "estimate" and "stderr", but this can also be set by the `col.estimate` and `col.stderr` arguments.
```{r simple-forest-plot, fig.width= 8, out.width = "90%"}
forest_plot(panels       = list(resultsA, resultsB),
            col.key      = "variable",
            exponentiate = TRUE)
```

### Using row labels
To use row labels, first create a data frame of labels. The data set should contain:

* a column used to match labels to the correct results, specified in the argument `col.key`.
* a column with labels for rows.
* further columns, as needed, for headings and subheadings.

If a particular level of label is not required it can be set to missing.

```{r rowlabels}
head(ckbplotr_row_labels)
```

In `forest_plot()` specify the `row.labels` argument as the data frame of labels you've created.

```{r using-row-labels, fig.height = 6}
forest_plot(panels            = list("Analysis A" = resultsA,
                                     "Analysis B" = resultsB),
            col.key           = "variable",
            row.labels        = ckbplotr_row_labels,
            exponentiate      = TRUE,
            xlim              = c(0.9, 1.1),
            xticks            = c(0.9, 1, 1.1),
            scalepoints       = TRUE,
            pointsize         = 3)
```

Use the `row.labels.levels` argument to choose which columns are used as headings/subheadings/labels. If not provided, all columns will be used.
For more details on customising forest plots, see ["More with forest plots"](forest_plot.html).



## Other plots
Use `+ ckb_style()` to covert an existing [ggplot2](https://ggplot2.tidyverse.org/) plot.
```{r a-plot, fig.height = 4, fig.width=8, out.width = "80%"}
plot <- ggplot(data = mpg,
               aes(x = displ, y = hwy)) +
  geom_point(size = 1)

ckbplot <- plot +
  ckb_style(xlims = c(0, 8),
            ylims = c(10, 50))

gridExtra::grid.arrange(plot, ckbplot, ncol = 2)
```

See ["Make any ggplot like a CKB plot"](plot_like_ckb.html) for more details.


## Confidence intervals

When standard errors are supplied to the `shape_plot()` and `forest_plot()` functions, confidence intervals are calculated as 95\% confidence intervals using the Normal approximation method (with critical value 1.96).

