---
title: "Making forest plots"
output:
  rmarkdown::html_vignette:
    fig_width: 7
    fig_height: 7
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{Making forest plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(ckbplotr)
```

## Different limits and ticks on each plot

The `make_forest_plot` function uses ggplot facets to place forest plots side-by-side. Facets cannot easily have different scales (or limits or ticks) applied, so it's not directly possible to have different limits and ticks on each forest plot.

However, one approach to work around this is to use `make_forest_plot` for each plot you need, remove the labels from all but the first, then arrange them side-by-side. The `gridExtra` package can be used for this last step.

Step 1: Use `make_forest_plot` for each plot.
```{r}
set.seed(57911624)
exampleresults <- function(){
  data.frame(variable = c('nmr_l_vldl_p', 'nmr_m_vldl_p', 'nmr_s_vldl_p',
                          'nmr_idl_p',
                          'nmr_l_ldl_p', 'nmr_m_ldl_p', 'nmr_s_ldl_p',
                          'nmr_l_vldl_tg', 'nmr_m_vldl_tg', 'nmr_s_vldl_tg',
                          'nmr_idl_tg',
                          'nmr_l_ldl_tg', 'nmr_m_ldl_tg', 'nmr_s_ldl_tg',
                          'nmr_l_vldl_c', 'nmr_m_vldl_c', 'nmr_s_vldl_c',
                          'nmr_idl_c',
                          'nmr_l_ldl_c', 'nmr_m_ldl_c', 'nmr_s_ldl_c'),
             estimate = rnorm(21, 0, 0.02),
             stderr   = 0.012 + abs(rnorm(21, 0, 0.015)),
             n        = round(runif(21, 100, 2000)),
             nb       = round(runif(21, 100, 2000)))
}
resultsA <- exampleresults()
resultsB <- exampleresults()
resultsC <- exampleresults()
resultsD <- exampleresults()
resultsE <- exampleresults()

headings <- data.frame(heading1 = rep(c("Lipoprotein particle concentration",
                                        "Triglycerides concentration",
                                        "Cholesterol concentration"), each = 7),
                       heading2 = rep(c("VLDL", "VLDL", "VLDL",
                                        "IDL",
                                        "LDL", "LDL", "LDL"), times = 3),
                       heading3 = rep(c("Large", "Medium", "Small",
                                        NA,
                                        "Large", "Medium", "Small") , times = 3),
                       variable = c('nmr_l_vldl_p', 'nmr_m_vldl_p', 'nmr_s_vldl_p',
                                    'nmr_idl_p',
                                    'nmr_l_ldl_p', 'nmr_m_ldl_p', 'nmr_s_ldl_p',
                                    'nmr_l_vldl_tg', 'nmr_m_vldl_tg', 'nmr_s_vldl_tg',
                                    'nmr_idl_tg',
                                    'nmr_l_ldl_tg', 'nmr_m_ldl_tg', 'nmr_s_ldl_tg',
                                    'nmr_l_vldl_c', 'nmr_m_vldl_c', 'nmr_s_vldl_c',
                                    'nmr_idl_c',
                                    'nmr_l_ldl_c', 'nmr_m_ldl_c', 'nmr_s_ldl_c'))


forestplot1 <- make_forest_plot(headings         = headings,
                                rows             = c("Lipoprotein particle concentration",
                                                     "Triglycerides concentration"),
                                cols             = list(resultsA),
                                exponentiate     = TRUE,
                                colnames         = c("Analysis A"),
                                col.key          = "variable",
                                ci.delim         = " - ",
                                xlim             = c(0.9, 1.1),
                                xticks           = c(0.9, 1, 1.1),
                                blankrows        = c(1, 1, 0, 1),
                                scalepoints      = TRUE,
                                pointsize        = 3,
                                col.left         = c("n"),
                                col.left.space   = c(0.02),
                                col.left.heading = c("No. of\nevents"),
                                col.right.space  = 0.02,
                                heading.space    = 2,
                                plot.space       = 8)

forestplot2 <- make_forest_plot(headings         = headings,
                                rows             = c("Lipoprotein particle concentration",
                                                     "Triglycerides concentration"),
                                cols             = list(resultsB),
                                exponentiate     = TRUE,
                                colnames         = c("Analysis B"),
                                col.key          = "variable",
                                ci.delim         = " - ",
                                xlim             = c(0.8, 1.2),
                                xticks           = c(0.8, 1, 1.2),
                                blankrows        = c(1, 1, 0, 1),
                                scalepoints      = TRUE,
                                pointsize        = 3,
                                col.left         = c("n"),
                                col.left.space   = c(0.02),
                                col.left.heading = c("No. of\nevents"),
                                col.right.space  = 0.02,
                                heading.space    = 2,
                                plot.space       = 8)
```

Step 2: Remove the axis text for all but the first plot.
```{r}
p1 <- forestplot1$plot

p2 <- forestplot2$plot + 
  theme(axis.text.y = element_blank())
```

Step 3: Arrange the plots using `gridExtra` (there may be other packages that also work). Adjust `widths` until the plots are the desired widths in your final output file.
```{r}
gridExtra::grid.arrange(p1, p2, ncol=2, widths = c(1,0.55))
```

Note that if `scalepoints = TRUE`, then this scaling is on a plot-by-plot basis so box sizes are not comparable between plots. However, if different axis scales are used then confidence intervals are not comparable either so this may be not be a problem.
