---
title: "Making shape plots"
output:
  rmarkdown::html_vignette:
    fig_width: 4
    fig_height: 4
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{Making shape plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 4,
  fig.height = 4,
  fig.align = "center"
)
```

```{r setup, include = FALSE}
library(ckbplotr)
```

The `make_shape_plot` function creates a plot of estimates and CIs against risk
factor levels using the `ggplot2` package.

The function returns a named list containing:

* plot: the plot
* code: ggplot2 code to generate the plot

The function uses the `plot_like_ckb` function to modify the theme and control
the look of the plot.

In RStudio, the ggplot2 code used to generate the plot will be shown in the 'Viewer'
pane. This code can be copied and further edited to adapt the plot.

## Prepare example data
```{r prep data}
results <- data.frame(
  est       = c(   1, 1.25,  1.5, 1.75,  1.7, 1.85, 2.25, 2.6),
  se        = c(0.08, 0.03, 0.07, 0.15, 0.08, 0.08, 0.07, 0.1),
  rf        = c(  20,   28,   38,   50, 18.5,   25,   37,  47),
  n         = c( 109,  103,  143,  104,  140,  134,  127,  99),
  is_female = c(   0,    0,    0,    0,    1,    1,    1,   1)
)
```

## Simple example
Given a data frame of estimates and standard errors (to be plotted on the y axis),
and risk factor levels (to be plotted on the x axis), a plot can be created.
```{r simple-example}
plot <- make_shape_plot(results[results$is_female == 0,],
                        col.x        = "rf",
                        col.estimate = "est",
                        col.stderr   = "se",
                        col.n        = "n",
                        xlims        = c(15, 50),
                        ylims        = c(0.5, 3),
                        scalepoints  = TRUE,
                        title        = NULL)
```

## Log scale
Set the `logscale` argument to true to use a log scale on the vertical axis.
```{r log-scale}
plot <- make_shape_plot(results[results$is_female == 0,],
                        col.x        = "rf",
                        col.estimate = "est",
                        col.stderr   = "se",
                        col.n        = "n",
                        xlims        = c(15, 50),
                        ylims        = c(0.7, 3),
                        ybreaks      = c(0.7, 1, 1.5, 2, 3),
                        scalepoints  = TRUE,
                        title        = NULL,
                        logscale     = TRUE)
```
If your estimates and standard errors are on the log scale (e.g. log hazard ratios), then set `exponentiate` to true. This will plot exp(estimates) and use a log scale for the axis (if `logscale` is not set).

## Using groups
The `col.group` argument can be supplied to plot results for different groups (using shades of grey for the fill colour). (Set `ciunder` to `TRUE` so that the confidence interval lines are hidden behind behind the point estimate squares.)
```{r groups}
plot <- make_shape_plot(results,
                        col.x        = "rf",
                        col.estimate = "est",
                        col.stderr   = "se",
                        col.n        = "n",
                        col.group    = "is_female",
                        xlims        = c(15,50),
                        ylims        = c(0.5, 3),
                        scalepoints  = TRUE,
                        title        = NULL,
                        ciunder      = TRUE)
```

## Adding lines
The `lines` argument will add lines (linear fit through estimates on plotted scale, weighted by inverse variance) for each group.
```{r lines}
plot <- make_shape_plot(results,
                        col.x        = "rf",
                        col.estimate = "est",
                        col.stderr   = "se",
                        col.n        = "n",
                        col.group    = "is_female",
                        xlims        = c(15,50),
                        ylims        = c(0.5, 3),
                        scalepoints  = TRUE,
                        title        = NULL,
                        ciunder      = TRUE,
                        lines        = TRUE)
```


## Setting aesthetics
The shape and fill colour of points, and colour of points and confidence interval lines can be set overall or on a per-point basis. This is done by setting arguments `shape`, `colour`, `cicolour`, `fill`, and `ciunder` to appropriate values, or to the name of a column containing values for each point.

The argument/columns, what they control, and the type:

| argument    | controls                                          | type      |
|-------------|---------------------------------------------------|-----------|
| shape       | plotting character for points                     | integer   |
| colour      | colour of points                                  | character |
| cicolour    | colour of CI lines                                | character |
| fill        | fill colour of points                             | character |
| ciunder     | if the CI line should be plotted before the point | logical   |


### Using values
If the argument doesn't match the anem of a column in the data, then the value will be used for all points.
```{r aesthetics with values}
plot <- make_shape_plot(results,
                        col.x        = "rf",
                        col.estimate = "est",
                        col.stderr   = "se",
                        col.n        = "n",
                        xlims        = c(15,50),
                        ylims        = c(0.5, 3),
                        scalepoints  = TRUE,
                        title        = NULL,
                        ciunder      = TRUE,
                        shape        = 23,
                        colour       = "black",
                        fill         = "red",
                        cicolour     = "blue")
```

### Using columns
If the argument matches a column name, then the values in the column will be used.
```{r aesthetics with columns}
results$fillcol <- "black"
results[results$is_female == 1,]$fillcol <- "orange"

plot <- make_shape_plot(results,
                        col.x        = "rf",
                        col.estimate = "est",
                        col.stderr   = "se",
                        col.n        = "n",
                        xlims        = c(15,50),
                        ylims        = c(0.5, 3),
                        scalepoints  = TRUE,
                        title        = NULL,
                        ciunder      = TRUE,
                        colour       = "fillcol",
                        fill         = "fillcol")
```

Specify a column for `cicolour` is useful for changing the colour of confidence interval lines that would otherwise be hidden.
```{r small ci}
results$cicol <- "black"
results[2,]$cicol <- "white"

plot <- make_shape_plot(results,
                        col.x        = "rf",
                        col.estimate = "est",
                        col.stderr   = "se",
                        col.n        = "n",
                        xlims        = c(15,50),
                        ylims        = c(0.5, 3),
                        scalepoints  = TRUE,
                        title        = NULL,
                        pointsize    = 7,
                        cicolour     = "cicol")
```

Note that the size of boxes depends on `pointsize`, but the true length of confidence interval lines depends on the size at which the plot is displayed or saved. So, first save your plot as an image then go back to change the colour of confidence interval lines as needed.
