---
title: "More with forest plots"
output:
  rmarkdown::html_vignette:
    fig_width: 7
    fig_height: 7
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{More with forest plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 7,
  fig.align = "center"
)
```

```{r setup, include = FALSE}
library(ckbplotr)
```


## Adding columns of text
```{r adding-columns-of-text}
resultsA <- dplyr::filter(ckbplotr_forest_data, name == "A")
resultsB <- dplyr::filter(ckbplotr_forest_data, name == "B")

forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = ckbplotr_row_labels,
                               row.labels.levels = c("heading", "subheading", "label"),
                               rows              = c("Triglycerides concentration",
                                                     "Lipoprotein particle concentration"),
                               exponentiate      = TRUE,
                               panel.headings    = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 1, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 3,
                               col.left          = c("n"),
                               col.left.heading  = c("No. of\nevents"),
                               col.heading.space = 1.5)
```


Multiple columns can be added by specifying vectors for `col.left`/`col.right` and `col.left.heading`/`col.right.heading`.
```{r adding-columns-of-text-2}
forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = ckbplotr_row_labels,
                               row.labels.levels = c("heading", "subheading", "label"),
                               rows              = c("Triglycerides concentration",
                                                     "Lipoprotein particle concentration"),
                               exponentiate      = TRUE,
                               panel.headings    = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 1, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 3,
                               col.left          = c("nb", "n"),
                               col.left.heading  = c("Some other\nnumbers", "No. of\nevents"),
                               col.heading.space = 1.5)
```

The `col.left.hjust` and `col.right.hjust` arguments set the horizontal justification of the columns (left = 0, center = 0.5, right = 1).

## Spacing
### Horizontal
The function attempts to set the positions of columns of text and spacing automatically. Where this does not produce a satisfactory layout, you can use the arguments `col.left.pos`, `col.right.pos`, `left.space`, `right.space`, and `mid.space` to control positions and spacing manually.

### Vertical
The plot will fill the vertical space available. Use `plot.margin` to change the top and bottom margins as needed.



## Setting colours and shapes, bold text and diamonds

The shape and fill colour of points, colour of points and confidence interval lines, bold text, and which estimates/CIs should be plotted as diamonds can be set overall or on a per-point basis. This is done by 
setting arguments `shape`, `colour`, `fill`, `ciunder`, `col.bold`, and `col.diamond` to appropriate values, or to the name of a column containing values for each point.

The argument/columns, what they control, and the type:

| argument    | controls                                          | type      |
|-------------|---------------------------------------------------|-----------|
| shape       | plotting character for points                     | integer   |
| colour      | colour of points and lines                        | character |
| fill        | fill colour of points                             | character |
| ciunder     | if the CI line should be plotted before the point | logical   |
| col.bold    | if text is bold                                   | logical   |
| col.diamond | if a diamond should be plotted                    | logical   |

Note that `col.bold`, and `col.diamond` must be column names in the supplied data frames, while the others can be fixed values or column names.

### Using values
```{r}
forestplot1 <- make_forest_plot(panels           = list(resultsA),
                                col.key          = "variable",
                                row.labels       = ckbplotr_row_labels,
                                row.labels.levels = c("heading", "subheading", "label"),
                                rows             = c("Triglycerides concentration"),
                                exponentiate     = TRUE,
                                panel.names      = c("Analysis A"),
                                ci.delim         = " - ",
                                xlim             = c(0.9, 1.1),
                                xticks           = c(0.9, 1, 1.1),
                                blankrows        = c(0, 1, 0, 1),
                                scalepoints      = TRUE,
                                pointsize        = 3,
                                col.left         = c("n"),
                                col.left.heading = c("No. of\nevents"),
                                shape            = 16,
                                colour           = "red",
                                cicolour         = "black",
                                ciunder          = TRUE)
```

### Using column names
```{r}
resultsA[9,"shape"] <- 16
resultsA[10, "bold"] <- TRUE
resultsA[11, "colour"] <- "red"
resultsA[12, "diamond"] <- TRUE
resultsA[13, "ciunder"] <- TRUE
resultsA[13, "shape"] <- 22
resultsA[13, "fill"] <- "white"

forestplot1 <- make_forest_plot(panels           = list(resultsA),
                                col.key          = "variable",
                                row.labels       = ckbplotr_row_labels,
                                row.labels.levels = c("heading", "subheading", "label"),
                                rows             = c("Triglycerides concentration"),
                                exponentiate     = TRUE,
                                panel.names      = c("Analysis A"),
                                ci.delim         = " - ",
                                xlim             = c(0.9, 1.1),
                                xticks           = c(0.9, 1, 1.1),
                                blankrows        = c(0, 1, 0, 1),
                                scalepoints      = TRUE,
                                pointsize        = 3,
                                col.left         = c("n"),
                                col.left.heading = c("No. of\nevents"),
                                shape            = "shape",
                                colour           = "colour",
                                col.bold         = "bold",
                                col.diamond      = "diamond",
                                ciunder          = "ciunder")
```

If a parameter is not set, then default values are used for these aesthetics. If a parameter is set, then every data frame provided in `cols` must contain a column with that name.

### Diamond shortcut
As an alternative to using `col.diamond`, provide a character vector in the `diamond` argument. In rows with these key values, estimates and CIs will be plotted using a diamond. (If a list is supplied, only the first element will be used.)
```{r}
forestplot1 <- make_forest_plot(panels           = list(resultsA),
                                col.key          = "variable",
                                row.labels       = ckbplotr_row_labels,
                                row.labels.levels = c("heading", "subheading", "label"),
                                rows             = c("Triglycerides concentration"),
                                exponentiate     = TRUE,
                                panel.names      = c("Analysis A"),
                                ci.delim         = " - ",
                                xlim             = c(0.9, 1.1),
                                xticks           = c(0.9, 1, 1.1),
                                blankrows        = c(0, 1, 0, 1),
                                scalepoints      = TRUE,
                                pointsize        = 3,
                                col.left         = c("n"),
                                col.left.heading = c("No. of\nevents"),
                                diamond          = c("nmr_l_ldl_tg", "nmr_m_ldl_tg"))
```




## Adding heterogeneity and trend test results and other text
The `addtext` argument can be used to add results of heterogeneity or trend test, or some other text, in the text column of estimates and CIs.

The automatic positioning of columns and spacing of panels does not take into account this additional text, so you may need to use the `right.space` and `col.right.pos` arguments for a satisfactory layout.

```{r, echo = FALSE}
resultsA_extra <- data.frame(variable   = c("nmr_s_ldl_p", "nmr_s_vldl_p"),
                             het_dof    = c("2",           NA),
                             het_stat   = c("12",          NA),
                             het_p      = c("=0.22",       NA),
                             trend_stat = c(NA,            "7"),
                             trend_p    = c(NA,            "=0.31"),
                             stringsAsFactors = FALSE)
resultsB_extra <- data.frame(variable   = c("nmr_s_ldl_p", "nmr_s_vldl_p"),
                             het_dof    = c("2",            NA),
                             het_stat   = c("14",           NA),
                             het_p      = c("=0.32",        NA),
                             trend_stat = c(NA,             "7"),
                             trend_p    = c(NA,             "=0.83"),
                             stringsAsFactors = FALSE)
```

```{r}
resultsA_extra
resultsB_extra
```

```{r, fig.width = 10}
forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = ckbplotr_row_labels,
                               row.labels.levels = c("heading", "subheading", "label"),
                               rows              = c("Lipoprotein particle concentration",
                                                     "Triglycerides concentration"),
                               exponentiate      = TRUE,
                               panel.headings    = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 0, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 3,
                               col.left          = c("n"),
                               col.left.heading  = c("No. of\nevents"),
                               col.heading.space = 1.5,
                               addtext           = list(resultsA_extra, resultsB_extra),
                               right.space = unit(35, "mm"))
```





















## Different limits and ticks on each plot

`make_forest_plot()` uses ggplot facets to place forest plots side-by-side. Facets cannot easily have different scales (or limits or ticks) applied, so it's not directly possible to have different limits and ticks on each forest plot.

However, one approach to work around this is to use `make_forest_plot()` for each plot you need, remove the labels from all but the first, then arrange them side-by-side. The `gridExtra` package can be used for this last step.

Step 1: Use `make_forest_plot()` for each plot.
```{r}
forestplot1 <- make_forest_plot(panels            = list(resultsA),
                                col.key           = "variable",
                                row.labels        = ckbplotr_row_labels,
                                row.labels.levels = c("heading", "subheading", "label"),
                                rows              = c("Lipoprotein particle concentration",
                                                      "Triglycerides concentration"),
                                exponentiate      = TRUE,
                                panel.names       = c("Analysis A"),
                                ci.delim          = " - ",
                                xlim              = c(0.9, 1.1),
                                xticks            = c(0.9, 1, 1.1),
                                blankrows         = c(1, 1, 0, 1),
                                scalepoints       = TRUE,
                                pointsize         = 3,
                                col.left          = c("n"),
                                col.left.heading  = c("No. of\nevents"),
                                col.heading.space = 1.5,
                                printplot         = FALSE)

forestplot2 <- make_forest_plot(panels            = list(resultsB),
                                col.key           = "variable",
                                row.labels        = ckbplotr_row_labels,
                                row.labels.levels = c("heading", "subheading", "label"),
                                rows              = c("Lipoprotein particle concentration",
                                                      "Triglycerides concentration"),
                                exponentiate      = TRUE,
                                panel.names       = c("Analysis B"),
                                ci.delim          = " - ",
                                xlim              = c(0.8, 1.2),
                                xticks            = c(0.8, 1, 1.2),
                                blankrows         = c(1, 1, 0, 1),
                                scalepoints       = TRUE,
                                pointsize         = 3,
                                col.left          = c("n"),
                                col.left.heading  = c("No. of\nevents"),
                                col.heading.space = 1.5,
                                printplot         = FALSE)
```

Step 2: Remove the axis text for all but the first plot.
```{r}
p1 <- forestplot1$plot

p2 <- forestplot2$plot + 
  theme(axis.text.y = element_blank())
```

Step 3: Arrange the plots using `gridExtra` (there may be other packages that also work). Adjust `widths` until the plots are the desired widths in your final output file. 
```{r}
gridExtra::grid.arrange(p1, p2, nrow = 1, widths = c(1, 0.52))
```

Note that if `scalepoints = TRUE` (and `minse` is not specified the same for each plot) then this scaling is on a plot-by-plot basis so box sizes are not comparable between plots. However, if different axis scales are used then confidence intervals are not comparable either so this may be not be a problem.



## Changing narrow confidence interval lines
When making a forest plot, if a confidence interval line is narrower than the point representing the estimate we may wish to plot the line differently. In this package, this is possible by using the `panel.width` argument in the `make_forest_plot()` function, then using the `fix_panel_width()` function.


### Changing the colour
To plot narrow confidence widths a different colour, set the `panel.width` argument of `make_forest_plot()` equal to the width each  panel will be in the final drawn output. The `cicolour` argument should also be a character vector - the last element will be used for narrow confidence intervals.

Then use the `fix_panel_width()` function, where the first argument is the plot just created and the second is the same panel width. This function returns a gtable object, which can be plotted using `grid::grid.draw` (or `plot()`). The panels widths will be fixed.

Notes: This has been designed to work well for shape 15 (the default) and 22. Use a small value for stroke (< 0.5) to avoid a wide border which is not accounted for. The calculation of sizes may not be entirely accurate, so check the plot and change the `plot.width` argument as needed. Confidence intervals are assumed to be centred on the point estimate.

```{r forest-plot}
forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = ckbplotr_row_labels,
                               row.labels.levels = c("heading", "subheading", "label"),
                               rows              = c("Lipoprotein particle concentration"),
                               exponentiate      = TRUE,
                               panel.names       = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 1, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 16,
                               col.left          = c("n"),
                               col.left.space    = c(0.02),
                               col.left.heading  = c("No. of\nevents"),
                               col.right.space   = 0.02,
                               col.heading.space = 0,
                               label.space       = 2,
                               panel.space       = 8,

                               # set panel width + CI colours
                               panel.width       = 25,
                               cicolour          = c("black", "white"),
                               printplot         = FALSE)

plot <- fix_panel_width(forestplot$plot, unit(25, "mm"))
grid::grid.draw(plot)
```

The `cicolour` argument can be a vector of all names of colours, or all names of columns (which contain colour names).
```{r forest-plot-2}

resultsA$cicol1 <- "red"
resultsB$cicol1 <- "blue"
resultsA$cicol2 <- "white"
resultsB$cicol2 <- "white"

forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = ckbplotr_row_labels,
                               row.labels.levels = c("heading", "subheading", "label"),
                               rows              = c("Lipoprotein particle concentration"),
                               exponentiate      = TRUE,
                               panel.names       = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 1, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 16,
                               col.left          = c("n"),
                               col.left.space    = c(0.02),
                               col.left.heading  = c("No. of\nevents"),
                               col.right.space   = 0.02,
                               col.heading.space = 0,
                               label.space       = 2,
                               panel.space       = 8,

                               # set panel width + CI colours
                               panel.width       = 25,
                               cicolour          = c("cicol1", "cicol2"),
                               printplot         = FALSE)

plot <- fix_panel_width(forestplot$plot, unit(25, "mm"))
grid::grid.draw(plot)
```


### Plotting lines under or over points
As well as changing colour, we may also wish to change if confidence interval lines are plotted under or over the point estimates. This can be done using the `plot.width` argument and using a logical (or character) vector for the `ciunder` argument.

```{r forest-plot-3}
forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = ckbplotr_row_labels,
                               row.labels.levels = c("heading", "subheading", "label"),
                               rows              = c("Lipoprotein particle concentration"),
                               exponentiate      = TRUE,
                               panel.names       = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 1, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 18,
                               col.left          = c("n"),
                               col.left.space    = c(0.02),
                               col.left.heading  = c("No. of\nevents"),
                               col.right.space   = 0.02,
                               col.heading.space = 0,
                               label.space       = 2,
                               panel.space       = 8,

                               # set panel width + CI under or over
                               panel.width       = 25,
                               shape             = 22,
                               stroke            = 0.5,
                               fill              = "white",
                               ciunder           = c(TRUE, FALSE),
                               printplot         = FALSE)

plot <- fix_panel_width(forestplot$plot, unit(25, "mm"))
grid::grid.draw(plot)
```

### Changing colour and plotting of lines

```{r forest-plot-4}
resultsA <- dplyr::select(resultsA, -ciunder)
resultsA$cicol1 <- "black"
resultsA$cicol2 <- "white"
resultsA$shape <- 15

resultsB <- dplyr::select(resultsB, -ciunder)
resultsB$cicol1 <- "black"
resultsB$cicol2 <- "black"
resultsB$shape <- 22

forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = ckbplotr_row_labels,
                               row.labels.levels = c("heading", "subheading", "label"),
                               rows              = c("Lipoprotein particle concentration"),
                               exponentiate      = TRUE,
                               panel.names       = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 1, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 18,
                               col.left          = c("n"),
                               col.left.space    = c(0.02),
                               col.left.heading  = c("No. of\nevents"),
                               col.right.space   = 0.02,
                               col.heading.space = 0,
                               label.space       = 2,
                               panel.space       = 8,

                               # set panel width + CI colours
                               panel.width       = 25,
                               shape             = "shape",
                               stroke            = 0.5,
                               cicolour          = c("cicol1", "cicol2"),
                               ciunder           = c(TRUE, FALSE),
                               fill              = "white",
                               printplot         = FALSE)

plot <- fix_panel_width(forestplot$plot, unit(25, "mm"))
grid::grid.draw(plot)
```




## Stroke
The `stroke` argument sets the stroke aesthetic for plotted shapes. See https://ggplot2.tidyverse.org/articles/ggplot2-specs.html for more details. The stroke size adds to total size of a shape, so unless `stroke = 0` the scaling of size by inverse variance will be slightly inaccurate (but there are probably more important things to worry about).
