---
title: "Making forest plots"
output:
  rmarkdown::html_vignette:
    fig_width: 7
    fig_height: 7
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{Making forest plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 7,
  fig.align = "center"
)
```

The `make_forest_plot` function creates a forest plot using the `ggplot2` graphics 
package. It is available as part of the `ckbplotr` package.

The function returns a named list containing:

* plot: the plot
* data: a data frame from which the plot is generated
* code: ggplot2 code to generate the plot

In RStudio the ggplot2 code used to generate the plot will be shown in the 
'Viewer' pane. If modifications are needed to the plot, then this code can be
copied, edited, and run as needed.


## Expected warning
You may get the following warning when using this function.

> Vectorized input to `element_text()` is not officially supported.
> Results may be unexpected or may change in future versions of ggplot2. 

This is related to how some headings used to label the rows of the forest plot are
made bold. This should not be a problem if the forest plot produced appears as expected.
(This hack is discussed [here](https://github.com/tidyverse/ggplot2/issues/3492).)



```{r setup, include = FALSE}
library(ckbplotr)
```

## Prepare data
First of all, do your analyses and put the results into data frames 
(one data frame for each forest plot column). Here we generate some example results.
```{r example forest plot results}
set.seed(57911624)
exampleresults <- function(){
  data.frame(variable = c('nmr_l_vldl_p', 'nmr_m_vldl_p', 'nmr_s_vldl_p',
                          'nmr_idl_p',
                          'nmr_l_ldl_p', 'nmr_m_ldl_p', 'nmr_s_ldl_p',
                          'nmr_l_vldl_tg', 'nmr_m_vldl_tg', 'nmr_s_vldl_tg',
                          'nmr_idl_tg',
                          'nmr_l_ldl_tg', 'nmr_m_ldl_tg', 'nmr_s_ldl_tg',
                          'nmr_l_vldl_c', 'nmr_m_vldl_c', 'nmr_s_vldl_c',
                          'nmr_idl_c',
                          'nmr_l_ldl_c', 'nmr_m_ldl_c', 'nmr_s_ldl_c'),
             estimate = rnorm(21, 0, 0.02),
             stderr   = 0.012 + abs(rnorm(21, 0, 0.015)),
             n        = round(runif(21, 100, 2000)),
             nb       = round(runif(21, 100, 2000)),
             shape    = rep(15, 21),
             colour   = "black",
             fill     = "black",
             bold     = FALSE,
             diamond  = FALSE,
             ciunder  = FALSE,
             stringsAsFactors = FALSE)
}
resultsA <- exampleresults()
resultsB <- exampleresults()
resultsC <- exampleresults()
resultsD <- exampleresults()
resultsE <- exampleresults()

resultsA[9,"shape"] <- 16
resultsA[10, "bold"] <- TRUE
resultsA[11, "colour"] <- "red"
resultsA[12, "diamond"] <- TRUE
resultsA[13, "ciunder"] <- TRUE
resultsA[13, "shape"] <- 22
resultsA[13, "fill"] <- "white"
```


## Simple forest plot
```{r make_forest_plot-example-1}
forestplot <- make_forest_plot(cols         = list(resultsA, resultsB),
                               exponentiate = TRUE,
                               colnames     = c("Analysis A", "Analysis B"),
                               col.key      = "variable")
```

## Forest plot with row headings
To use headings, create a data frame of headings and specify arguments `headings` and
`rows`.

This data frame must contain columns 'heading1', 'heading2' and 'heading3',
which can be set to NA if not needed for a particular row. For example, if you 
do not need a third level of heading for a particular row, then set `heading3=NA`
for that row. If you only need one level of headings, then set `heading2=NA` and
`heading3=NA` for every row.

The data set must also contain a column with the same name as a column in each 
of the results data sets. This column is used to match headings to the correct 
results, and is specified in the argument `col.key`.

The argument `rows` specifies which results should be included in the plot, by
giving some headings from the heading1 column of the headings data frame.

(The order given in the `rows` argument decides the order in which the heading1 
labels will be in the plot (top to bottom). The heading2 and heading3 labels
will be in the order that they are in the headings data frame.)

```{r forest plot headings, eval=TRUE, echo=FALSE}
# Example headings for every row
# This could be loaded from a csv containing headings that you use often
headings <- data.frame(heading1 = rep(c("Lipoprotein particle concentration",
                                        "Triglycerides concentration",
                                        "Cholesterol concentration"), each = 7),
                       heading2 = rep(c("VLDL", "VLDL", "VLDL",
                                        "IDL",
                                        "LDL", "LDL", "LDL"), times = 3),
                       heading3 = rep(c("Large", "Medium", "Small",
                                        NA,
                                        "Large", "Medium", "Small") , times = 3),
                       variable = c('nmr_l_vldl_p', 'nmr_m_vldl_p', 'nmr_s_vldl_p',
                                    'nmr_idl_p',
                                    'nmr_l_ldl_p', 'nmr_m_ldl_p', 'nmr_s_ldl_p',
                                    'nmr_l_vldl_tg', 'nmr_m_vldl_tg', 'nmr_s_vldl_tg',
                                    'nmr_idl_tg',
                                    'nmr_l_ldl_tg', 'nmr_m_ldl_tg', 'nmr_s_ldl_tg',
                                    'nmr_l_vldl_c', 'nmr_m_vldl_c', 'nmr_s_vldl_c',
                                    'nmr_idl_c',
                                    'nmr_l_ldl_c', 'nmr_m_ldl_c', 'nmr_s_ldl_c'),
                       stringsAsFactors = FALSE)
headings
```

```{r make_forest_plot-example-2}
forestplot <- make_forest_plot(headings         = headings,
                               rows             = c("Lipoprotein particle concentration",
                                                    "Triglycerides concentration"),
                               cols             = list(resultsA, resultsB),
                               exponentiate     = TRUE,
                               colnames         = c("Analysis A", "Analysis B"),
                               col.key          = "variable",
                               ci.delim         = " - ",
                               xlim             = c(0.9, 1.1),
                               xticks           = c(0.9, 1, 1.1),
                               blankrows        = c(1, 1, 0, 1),
                               scalepoints      = TRUE,
                               pointsize        = 3,
                               col.left         = c("n"),
                               col.left.space   = c(0.02),
                               col.left.heading = c("No. of\nevents"),
                               col.right.space  = 0.02,
                               col.heading.space = 1.5,
                               heading.space    = 2,
                               plot.space       = 8)
```


## Setting colours and shapes, bold text and diamonds

The shape and fill colour of points, colour of points and confidence interval lines, bold text, and which estimates/CIs should be plotted as diamonds can be set overall or on a per-point basis. This is done by 
setting arguments `shape`, `colour`, `fill`, `ciunder`, `col.bold`, and `col.diamond` to appropriate values, or to the name of a column containing values for each point.

The argument/columns, what they control, and the type:

| argument    | controls                                          | type      |
|-------------|---------------------------------------------------|-----------|
| shape       | plotting character for points                     | integer   |
| colour      | colour of points and lines                        | character |
| fill        | fill colour of points                             | character |
| ciunder     | if the CI line should be plotted before the point | logical   |
| col.bold    | if text is bold                                   | logical   |
| col.diamond | if a diamond should be plotted                    | logical   |

Note that `col.bold`, and `col.diamond` must be column names in the supplied data frames, while the others can be fixed values or column names.

### Using values
```{r}
forestplot1 <- make_forest_plot(headings         = headings,
                                rows             = c("Triglycerides concentration"),
                                cols             = list(resultsA),
                                exponentiate     = TRUE,
                                colnames         = c("Analysis A"),
                                col.key          = "variable",
                                ci.delim         = " - ",
                                xlim             = c(0.9, 1.1),
                                xticks           = c(0.9, 1, 1.1),
                                blankrows        = c(0, 1, 0, 1),
                                scalepoints      = TRUE,
                                pointsize        = 3,
                                col.left         = c("n"),
                                col.left.space   = c(0.02),
                                col.left.heading = c("No. of\nevents"),
                                col.right.space  = 0.02,
                                heading.space    = 2,
                                plot.space       = 8,
                                shape            = 16,
                                colour           = "red",
                                cicolour         = "black",
                                ciunder          = TRUE)
```

### Using column names
```{r}
forestplot1 <- make_forest_plot(headings         = headings,
                                rows             = c("Triglycerides concentration"),
                                cols             = list(resultsA),
                                exponentiate     = TRUE,
                                colnames         = c("Analysis A"),
                                col.key          = "variable",
                                ci.delim         = " - ",
                                xlim             = c(0.9, 1.1),
                                xticks           = c(0.9, 1, 1.1),
                                blankrows        = c(0, 1, 0, 1),
                                scalepoints      = TRUE,
                                pointsize        = 3,
                                col.left         = c("n"),
                                col.left.space   = c(0.02),
                                col.left.heading = c("No. of\nevents"),
                                col.right.space  = 0.02,
                                heading.space    = 2,
                                plot.space       = 8,
                                shape            = "shape",
                                colour           = "colour",
                                col.bold         = "bold",
                                col.diamond      = "diamond",
                                ciunder          = "ciunder")
```

If a parameter is not set, then default values are used for these aesthetics. If a parameter is set, then every data frame provided in `cols` must contain a column with that name.

### Diamond shortcut
As an alternative to using `col.diamond`, provide a character vector in the `diamond` argument. In rows with these key values, estimates and CIs will be plotted using a diamond. (If a list is supplied, only the first element will be used.)
```{r}
forestplot1 <- make_forest_plot(headings         = headings,
                                rows             = c("Triglycerides concentration"),
                                cols             = list(resultsA),
                                exponentiate     = TRUE,
                                colnames         = c("Analysis A"),
                                col.key          = "variable",
                                ci.delim         = " - ",
                                xlim             = c(0.9, 1.1),
                                xticks           = c(0.9, 1, 1.1),
                                blankrows        = c(0, 1, 0, 1),
                                scalepoints      = TRUE,
                                pointsize        = 3,
                                col.left         = c("n"),
                                col.left.space   = c(0.02),
                                col.left.heading = c("No. of\nevents"),
                                col.right.space  = 0.02,
                                heading.space    = 2,
                                plot.space       = 8,
                                diamond          = c("nmr_l_ldl_tg", "nmr_m_ldl_tg"))
```




## Adding heterogeneity and trend test results and other text
The `addtext` argument can be used to add results of heterogeneity or trend test, or some other text, in the text column of estimates and CIs.

You will probably need to increase `plot.space` and widen the right margin as shown to stop overlapping text or overrunning the edge of the plot.
```{r, fig.width = 10}
resultsA_extra <- data.frame(variable   = c("nmr_s_ldl_p", "nmr_s_vldl_p"),
                             het_dof    = c("2",           NA),
                             het_stat   = c("12",          NA),
                             het_p      = c("=0.22",       NA),
                             trend_stat = c(NA,            "7"),
                             trend_p    = c(NA,            "=0.31"),
                             stringsAsFactors = FALSE)
resultsB_extra <- data.frame(variable   = c("nmr_s_ldl_p", "nmr_s_vldl_p"),
                             het_dof    = c("2",            NA),
                             het_stat   = c("14",           NA),
                             het_p      = c("=0.32",        NA),
                             trend_stat = c(NA,             "7"),
                             trend_p    = c(NA,             "=0.83"),
                             stringsAsFactors = FALSE)

forestplot <- make_forest_plot(headings         = headings,
                               rows             = c("Lipoprotein particle concentration",
                                                    "Triglycerides concentration"),
                               cols             = list(resultsA, resultsB),
                               exponentiate     = TRUE,
                               colnames         = c("Analysis A", "Analysis B"),
                               col.key          = "variable",
                               ci.delim         = " - ",
                               xlim             = c(0.9, 1.1),
                               xticks           = c(0.9, 1, 1.1),
                               blankrows        = c(1, 0, 0, 1),
                               scalepoints      = TRUE,
                               pointsize        = 3,
                               col.left         = c("n"),
                               col.left.space   = c(0.02),
                               col.left.heading = c("No. of\nevents"),
                               col.right.space  = 0.02,
                               col.heading.space = 1.5,
                               heading.space    = 2,
                               plot.space       = 10,
                               addtext          = list(resultsA_extra, resultsB_extra),
                               printplot        = FALSE)

forestplot$plot + theme(plot.margin = unit(c(2,10,2,0), "lines"))
```





















## Different limits and ticks on each plot

The `make_forest_plot` function uses ggplot facets to place forest plots side-by-side. Facets cannot easily have different scales (or limits or ticks) applied, so it's not directly possible to have different limits and ticks on each forest plot.

However, one approach to work around this is to use `make_forest_plot` for each plot you need, remove the labels from all but the first, then arrange them side-by-side. The `gridExtra` package can be used for this last step.

Step 1: Use `make_forest_plot` for each plot.
```{r}
forestplot1 <- make_forest_plot(headings         = headings,
                                rows             = c("Lipoprotein particle concentration",
                                                     "Triglycerides concentration"),
                                cols             = list(resultsA),
                                exponentiate     = TRUE,
                                colnames         = c("Analysis A"),
                                col.key          = "variable",
                                ci.delim         = " - ",
                                xlim             = c(0.9, 1.1),
                                xticks           = c(0.9, 1, 1.1),
                                blankrows        = c(1, 1, 0, 1),
                                scalepoints      = TRUE,
                                pointsize        = 3,
                                col.left         = c("n"),
                                col.left.space   = c(0.02),
                                col.left.heading = c("No. of\nevents"),
                                col.right.space  = 0.02,
                                col.heading.space = 1.5,
                                heading.space    = 2,
                                plot.space       = 8,
                                printplot        = FALSE)

forestplot2 <- make_forest_plot(headings         = headings,
                                rows             = c("Lipoprotein particle concentration",
                                                     "Triglycerides concentration"),
                                cols             = list(resultsB),
                                exponentiate     = TRUE,
                                colnames         = c("Analysis B"),
                                col.key          = "variable",
                                ci.delim         = " - ",
                                xlim             = c(0.8, 1.2),
                                xticks           = c(0.8, 1, 1.2),
                                blankrows        = c(1, 1, 0, 1),
                                scalepoints      = TRUE,
                                pointsize        = 3,
                                col.left         = c("n"),
                                col.left.space   = c(0.02),
                                col.left.heading = c("No. of\nevents"),
                                col.right.space  = 0.02,
                                col.heading.space = 1.5,
                                heading.space    = 2,
                                plot.space       = 8,
                                printplot        = FALSE)
```

Step 2: Remove the axis text for all but the first plot.
```{r}
p1 <- forestplot1$plot

p2 <- forestplot2$plot + 
  theme(axis.text.y = element_blank())
```

Step 3: Arrange the plots using `gridExtra` (there may be other packages that also work). Adjust `widths` until the plots are the desired widths in your final output file.
```{r}
gridExtra::grid.arrange(p1, p2, ncol=2, widths = c(1,0.55))
```

Note that if `scalepoints = TRUE`, then this scaling is on a plot-by-plot basis so box sizes are not comparable between plots. However, if different axis scales are used then confidence intervals are not comparable either so this may be not be a problem.

## Stroke
The `stroke` argument sets the stroke aesthetic for plotted shapes. See https://ggplot2.tidyverse.org/articles/ggplot2-specs.html for more details. The stroke size adds to total size of a shape, so unless `stroke = 0` the scaling of size by inverse variance will be slightly inaccurate (but there are probably more important things to worry about).
