---
title: "More with forest plots"
output:
  rmarkdown::html_vignette:
    fig_width: 7
    fig_height: 7
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{More with forest plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 7,
  fig.align = "center"
)
```

```{r setup, include = FALSE}
library(ckbplotr)

set.seed(57911624)
exampleresults <- function(){
  data.frame(variable = c('nmr_l_vldl_p', 'nmr_m_vldl_p', 'nmr_s_vldl_p',
                          'nmr_idl_p',
                          'nmr_l_ldl_p', 'nmr_m_ldl_p', 'nmr_s_ldl_p',
                          'nmr_l_vldl_tg', 'nmr_m_vldl_tg', 'nmr_s_vldl_tg',
                          'nmr_idl_tg',
                          'nmr_l_ldl_tg', 'nmr_m_ldl_tg', 'nmr_s_ldl_tg',
                          'nmr_l_vldl_c', 'nmr_m_vldl_c', 'nmr_s_vldl_c',
                          'nmr_idl_c',
                          'nmr_l_ldl_c', 'nmr_m_ldl_c', 'nmr_s_ldl_c'),
             estimate = rnorm(21, 0, 0.02),
             stderr   = 0.012 + abs(rnorm(21, 0, 0.015)),
             n        = round(runif(21, 100, 2000)),
             nb       = round(runif(21, 100, 2000)),
             shape    = rep(15, 21),
             colour   = "black",
             fill     = "black",
             bold     = FALSE,
             diamond  = FALSE,
             ciunder  = FALSE,
             stringsAsFactors = FALSE)
}
resultsA <- exampleresults()
resultsB <- exampleresults()
resultsC <- exampleresults()
resultsD <- exampleresults()
resultsE <- exampleresults()

mylabels <- data.frame(heading1 = rep(c("Lipoprotein particle concentration",
                                        "Triglycerides concentration",
                                        "Cholesterol concentration"), each = 7),
                       heading2 = rep(c("VLDL", "VLDL", "VLDL",
                                        "IDL",
                                        "LDL", "LDL", "LDL"), times = 3),
                       heading3 = rep(c("Large", "Medium", "Small",
                                        NA,
                                        "Large", "Medium", "Small") , times = 3),
                       variable = c('nmr_l_vldl_p', 'nmr_m_vldl_p', 'nmr_s_vldl_p',
                                    'nmr_idl_p',
                                    'nmr_l_ldl_p', 'nmr_m_ldl_p', 'nmr_s_ldl_p',
                                    'nmr_l_vldl_tg', 'nmr_m_vldl_tg', 'nmr_s_vldl_tg',
                                    'nmr_idl_tg',
                                    'nmr_l_ldl_tg', 'nmr_m_ldl_tg', 'nmr_s_ldl_tg',
                                    'nmr_l_vldl_c', 'nmr_m_vldl_c', 'nmr_s_vldl_c',
                                    'nmr_idl_c',
                                    'nmr_l_ldl_c', 'nmr_m_ldl_c', 'nmr_s_ldl_c'),
                       stringsAsFactors = FALSE)
```


## Adding columns of text
```{r adding-columns-of-text}
forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = mylabels,
                               rows              = c("Triglycerides concentration",
                                                     "Lipoprotein particle concentration"),
                               exponentiate      = TRUE,
                               panel.headings    = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 1, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 3,
                               col.left          = c("n"),
                               col.left.heading  = c("No. of\nevents"),
                               col.heading.space = 1.5)
```


Multiple columns can be added by specifying vectors for `col.left`/`col.right` and `col.left.heading`/`col.right.heading`. Make sure to also use a vector for `col.left.space`/`col.right.space` so that the columns do not overlap.
```{r adding-columns-of-text-2}
forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = mylabels,
                               rows              = c("Triglycerides concentration",
                                                     "Lipoprotein particle concentration"),
                               exponentiate      = TRUE,
                               panel.headings    = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 1, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 3,
                               col.left          = c("nb", "n"),
                               col.left.heading  = c("Some other\nnumbers", "No. of\nevents"),
                               col.left.space    = c(0.6, 0.02),
                               col.heading.space = 1.5)
```

The `col.left.hjust` and `col.right.hjust` arguments set the horizontal justification of the columns (left = 0, center = 0.5, right = 1).

## Spacing
### Horizontal
Controlling horizontal spacing when using `make_forest_plot()` can be tedious. The suggested workflow is:

1. Set `plot.margin`. This is a numerical vector of length four. The 2nd element controls the margin to the right of the plot, and the 4th element controls the margin to the left of the plot (the 1st and 3rd are the top and bottom margins). The unit is set by the `units` argument which has default "lines".
2. Set `left.space` and `right.space` to allow space to the left and right of each panel, for example for columns of text. The unit is given by the `units` argument.
3. Set `mid.space` to add additional space between panels.
4. Create your plot and output it to the final file. For example, use `ggplot2::ggsave()` to save it as a PDF. The x-axes of the panels will fill the remaing horizontal space available.
5. Now use the `col.left.space` and `col.right.space` arguments to position the your `col.left` and `col.right` columns. A value of 1 is equal to the width of the x-axis. If there is not enough space for your columns, go back to step 2.

For a visual guide to this workflow, see [https://github.com/neilstats/ckbplotr-help/make_forest_plot_hspacing.pdf](https://github.com/neilstats/ckbplotr-help/make_forest_plot_hspacing.pdf).


### Vertical
The plot will fill the vertical space available. Use `plot.margin` to change the top and bottom margins as needed.



## Setting colours and shapes, bold text and diamonds

The shape and fill colour of points, colour of points and confidence interval lines, bold text, and which estimates/CIs should be plotted as diamonds can be set overall or on a per-point basis. This is done by 
setting arguments `shape`, `colour`, `fill`, `ciunder`, `col.bold`, and `col.diamond` to appropriate values, or to the name of a column containing values for each point.

The argument/columns, what they control, and the type:

| argument    | controls                                          | type      |
|-------------|---------------------------------------------------|-----------|
| shape       | plotting character for points                     | integer   |
| colour      | colour of points and lines                        | character |
| fill        | fill colour of points                             | character |
| ciunder     | if the CI line should be plotted before the point | logical   |
| col.bold    | if text is bold                                   | logical   |
| col.diamond | if a diamond should be plotted                    | logical   |

Note that `col.bold`, and `col.diamond` must be column names in the supplied data frames, while the others can be fixed values or column names.

### Using values
```{r}
resultsA[9,"shape"] <- 16
resultsA[10, "bold"] <- TRUE
resultsA[11, "colour"] <- "red"
resultsA[12, "diamond"] <- TRUE
resultsA[13, "ciunder"] <- TRUE
resultsA[13, "shape"] <- 22
resultsA[13, "fill"] <- "white"

forestplot1 <- make_forest_plot(panels           = list(resultsA),
                                col.key          = "variable",
                                row.labels       = mylabels,
                                rows             = c("Triglycerides concentration"),
                                exponentiate     = TRUE,
                                panel.names      = c("Analysis A"),
                                ci.delim         = " - ",
                                xlim             = c(0.9, 1.1),
                                xticks           = c(0.9, 1, 1.1),
                                blankrows        = c(0, 1, 0, 1),
                                scalepoints      = TRUE,
                                pointsize        = 3,
                                col.left         = c("n"),
                                col.left.heading = c("No. of\nevents"),
                                shape            = 16,
                                colour           = "red",
                                cicolour         = "black",
                                ciunder          = TRUE)
```

### Using column names
```{r}
forestplot1 <- make_forest_plot(panels           = list(resultsA),
                                col.key          = "variable",
                                row.labels       = mylabels,
                                rows             = c("Triglycerides concentration"),
                                exponentiate     = TRUE,
                                panel.names      = c("Analysis A"),
                                ci.delim         = " - ",
                                xlim             = c(0.9, 1.1),
                                xticks           = c(0.9, 1, 1.1),
                                blankrows        = c(0, 1, 0, 1),
                                scalepoints      = TRUE,
                                pointsize        = 3,
                                col.left         = c("n"),
                                col.left.heading = c("No. of\nevents"),
                                shape            = "shape",
                                colour           = "colour",
                                col.bold         = "bold",
                                col.diamond      = "diamond",
                                ciunder          = "ciunder")
```

If a parameter is not set, then default values are used for these aesthetics. If a parameter is set, then every data frame provided in `cols` must contain a column with that name.

### Diamond shortcut
As an alternative to using `col.diamond`, provide a character vector in the `diamond` argument. In rows with these key values, estimates and CIs will be plotted using a diamond. (If a list is supplied, only the first element will be used.)
```{r}
forestplot1 <- make_forest_plot(panels           = list(resultsA),
                                col.key          = "variable",
                                row.labels       = mylabels,
                                rows             = c("Triglycerides concentration"),
                                exponentiate     = TRUE,
                                panel.names      = c("Analysis A"),
                                ci.delim         = " - ",
                                xlim             = c(0.9, 1.1),
                                xticks           = c(0.9, 1, 1.1),
                                blankrows        = c(0, 1, 0, 1),
                                scalepoints      = TRUE,
                                pointsize        = 3,
                                col.left         = c("n"),
                                col.left.heading = c("No. of\nevents"),
                                diamond          = c("nmr_l_ldl_tg", "nmr_m_ldl_tg"))
```




## Adding heterogeneity and trend test results and other text
The `addtext` argument can be used to add results of heterogeneity or trend test, or some other text, in the text column of estimates and CIs.

```{r, echo = FALSE}
resultsA_extra <- data.frame(variable   = c("nmr_s_ldl_p", "nmr_s_vldl_p"),
                             het_dof    = c("2",           NA),
                             het_stat   = c("12",          NA),
                             het_p      = c("=0.22",       NA),
                             trend_stat = c(NA,            "7"),
                             trend_p    = c(NA,            "=0.31"),
                             stringsAsFactors = FALSE)
resultsB_extra <- data.frame(variable   = c("nmr_s_ldl_p", "nmr_s_vldl_p"),
                             het_dof    = c("2",            NA),
                             het_stat   = c("14",           NA),
                             het_p      = c("=0.32",        NA),
                             trend_stat = c(NA,             "7"),
                             trend_p    = c(NA,             "=0.83"),
                             stringsAsFactors = FALSE)
```

```{r}
resultsA_extra
resultsB_extra
```

```{r, fig.width = 10}
forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = mylabels,
                               rows              = c("Lipoprotein particle concentration",
                                                     "Triglycerides concentration"),
                               exponentiate      = TRUE,
                               panel.headings    = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 0, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 3,
                               col.left          = c("n"),
                               col.left.heading  = c("No. of\nevents"),
                               col.heading.space = 1.5,
                               addtext           = list(resultsA_extra, resultsB_extra),
                               printplot         = FALSE)

forestplot$plot + theme(plot.margin = unit(c(2,10,2,1), "lines"))
```





















## Different limits and ticks on each plot

`make_forest_plot()` uses ggplot facets to place forest plots side-by-side. Facets cannot easily have different scales (or limits or ticks) applied, so it's not directly possible to have different limits and ticks on each forest plot.

However, one approach to work around this is to use `make_forest_plot()` for each plot you need, remove the labels from all but the first, then arrange them side-by-side. The `gridExtra` package can be used for this last step.

Step 1: Use `make_forest_plot()` for each plot.
```{r}
forestplot1 <- make_forest_plot(panels            = list(resultsA),
                                col.key           = "variable",
                                row.labels        = mylabels,
                                rows              = c("Lipoprotein particle concentration",
                                                      "Triglycerides concentration"),
                                exponentiate      = TRUE,
                                panel.names       = c("Analysis A"),
                                ci.delim          = " - ",
                                xlim              = c(0.9, 1.1),
                                xticks            = c(0.9, 1, 1.1),
                                blankrows         = c(1, 1, 0, 1),
                                scalepoints       = TRUE,
                                pointsize         = 3,
                                col.left          = c("n"),
                                col.left.heading  = c("No. of\nevents"),
                                col.heading.space = 1.5,
                                printplot         = FALSE)

forestplot2 <- make_forest_plot(panels            = list(resultsB),
                                col.key           = "variable",
                                row.labels        = mylabels,
                                rows              = c("Lipoprotein particle concentration",
                                                      "Triglycerides concentration"),
                                exponentiate      = TRUE,
                                panel.names       = c("Analysis B"),
                                ci.delim          = " - ",
                                xlim              = c(0.8, 1.2),
                                xticks            = c(0.8, 1, 1.2),
                                blankrows         = c(1, 1, 0, 1),
                                scalepoints       = TRUE,
                                pointsize         = 3,
                                col.left          = c("n"),
                                col.left.heading  = c("No. of\nevents"),
                                col.heading.space = 1.5,
                                printplot         = FALSE)
```

Step 2: Remove the axis text for all but the first plot.
```{r}
p1 <- forestplot1$plot

p2 <- forestplot2$plot + 
  theme(axis.text.y = element_blank())
```

Step 3: Arrange the plots using `gridExtra` (there may be other packages that also work). Adjust `widths` until the plots are the desired widths in your final output file. 
```{r}
gridExtra::grid.arrange(p1, p2, nrow = 1, widths = c(1, 0.52))
```

Note that if `scalepoints = TRUE` (and `minse` is not specified the same for each plot) then this scaling is on a plot-by-plot basis so box sizes are not comparable between plots. However, if different axis scales are used then confidence intervals are not comparable either so this may be not be a problem.



## Changing narrow confidence interval lines
When making a forest plot, if a confidence interval line is narrower than the point representing the estimate we may wish to plot the line differently. In this package, this is possible by using the `panel.width` argument in the `make_forest_plot()` function, then using the `fix_panel_width()` function.


### Changing the colour
To plot narrow confidence widths a different colour, set the `panel.width` argument of `make_forest_plot()` equal to the width (in mm) each  panel will be in the final drawn output. The `cicolour` argument should also be a character vector - the last element will be used for narrow confidence intervals.

Then use the `fix_panel_width()` function, where the first argument is the plot just created and the second is the same panel width. This function returns a gtable object, which can be plotted using `grid::grid.draw` (or `plot()`). The panels widths will be fixed.

Notes: This has been designed to work well for shape 15 (the default) and 22. Use a small value for stroke (< 0.5) to avoid a wide border which is not accounted for. The calculation of sizes may not be entirely accurate, so check the plot and change the `plot.width` argument as needed. Confidence intervals are assumed to be centred on the point estimate.

```{r forest-plot}
forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = mylabels,
                               rows              = c("Lipoprotein particle concentration"),
                               exponentiate      = TRUE,
                               panel.names       = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 1, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 16,
                               col.left          = c("n"),
                               col.left.space    = c(0.02),
                               col.left.heading  = c("No. of\nevents"),
                               col.right.space   = 0.02,
                               col.heading.space = 0,
                               label.space       = 2,
                               panel.space       = 8,

                               # set panel width + CI colours
                               panel.width       = 25,
                               cicolour          = c("black", "white"),
                               printplot         = FALSE)

plot <- fix_panel_width(forestplot$plot, 25)
grid::grid.draw(plot)
```

The `cicolour` argument can be a vector of all names of colours, or all names of columns (which contain colour names).
```{r forest-plot-2}

resultsA$cicol1 <- "red"
resultsB$cicol1 <- "blue"
resultsA$cicol2 <- "white"
resultsB$cicol2 <- "white"

forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = mylabels,
                               rows              = c("Lipoprotein particle concentration"),
                               exponentiate      = TRUE,
                               panel.names       = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 1, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 16,
                               col.left          = c("n"),
                               col.left.space    = c(0.02),
                               col.left.heading  = c("No. of\nevents"),
                               col.right.space   = 0.02,
                               col.heading.space = 0,
                               label.space       = 2,
                               panel.space       = 8,

                               # set panel width + CI colours
                               panel.width       = 25,
                               cicolour          = c("cicol1", "cicol2"),
                               printplot         = FALSE)

plot <- fix_panel_width(forestplot$plot, 25)
grid::grid.draw(plot)
```


### Plotting lines under or over points
As well as changing colour, we may also wish to change if confidence interval lines are plotted under or over the point estimates. This can be done using the `plot.width` argument and using a logical (or character) vector for the `ciunder` argument.

```{r forest-plot-3}
forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = mylabels,
                               rows              = c("Lipoprotein particle concentration"),
                               exponentiate      = TRUE,
                               panel.names       = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 1, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 18,
                               col.left          = c("n"),
                               col.left.space    = c(0.02),
                               col.left.heading  = c("No. of\nevents"),
                               col.right.space   = 0.02,
                               col.heading.space = 0,
                               label.space       = 2,
                               panel.space       = 8,

                               # set panel width + CI under or over
                               panel.width       = 25,
                               shape             = 22,
                               stroke            = 0.5,
                               fill              = "white",
                               ciunder           = c(TRUE, FALSE),
                               printplot         = FALSE)

plot <- fix_panel_width(forestplot$plot, 25)
grid::grid.draw(plot)
```

### Changing colour and plotting of lines

```{r forest-plot-4}
resultsA <- dplyr::select(resultsA, -ciunder)
resultsA$cicol1 <- "black"
resultsA$cicol2 <- "white"
resultsA$shape <- 15

resultsB <- dplyr::select(resultsB, -ciunder)
resultsB$cicol1 <- "black"
resultsB$cicol2 <- "black"
resultsB$shape <- 22

forestplot <- make_forest_plot(panels            = list(resultsA, resultsB),
                               col.key           = "variable",
                               row.labels        = mylabels,
                               rows              = c("Lipoprotein particle concentration"),
                               exponentiate      = TRUE,
                               panel.names       = c("Analysis A", "Analysis B"),
                               ci.delim          = " - ",
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               blankrows         = c(1, 1, 0, 1),
                               scalepoints       = TRUE,
                               pointsize         = 18,
                               col.left          = c("n"),
                               col.left.space    = c(0.02),
                               col.left.heading  = c("No. of\nevents"),
                               col.right.space   = 0.02,
                               col.heading.space = 0,
                               label.space       = 2,
                               panel.space       = 8,

                               # set panel width + CI colours
                               panel.width       = 25,
                               shape             = "shape",
                               stroke            = 0.5,
                               cicolour          = c("cicol1", "cicol2"),
                               ciunder           = c(TRUE, FALSE),
                               fill              = "white",
                               printplot         = FALSE)

plot <- fix_panel_width(forestplot$plot, 25)
grid::grid.draw(plot)
```




## Stroke
The `stroke` argument sets the stroke aesthetic for plotted shapes. See https://ggplot2.tidyverse.org/articles/ggplot2-specs.html for more details. The stroke size adds to total size of a shape, so unless `stroke = 0` the scaling of size by inverse variance will be slightly inaccurate (but there are probably more important things to worry about).
