---
title: "Page layouts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Page layouts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The package does not currently include any functions for composing multiple plots, but these package may be useful:

* grid and [gridExtra](https://cran.r-project.org/package=gridExtra)
* [patchwork](https://patchwork.data-imaginist.com/)
* [cowplot](https://cran.r-project.org/package=cowplot)
* [egg](https://cran.r-project.org/package=egg)
* [multipanelfigure](https://cran.r-project.org/package=multipanelfigure)



## Example
Using the [grid](https://stat.ethz.ch/R-manual/R-devel/library/grid/html/00Index.html) and [gridExtra](https://cran.r-project.org/package=gridExtra) packages gives a lot of control but may need more work to create the desired output. Here is an example of placing two plots, with a title and footer, on an A4 page in a PDF document.
```{r, eval = F}
# Make the plots, with suitable margins
shapeplot <- make_shape_plot(ckbplotr_shape_data[ckbplotr_shape_data$is_female == 0,],
                        col.x        = "rf",
                        col.estimate = "est",
                        col.stderr   = "se",
                        col.n        = "n",
                        xlims        = c(15, 50),
                        ylims        = c(0.5, 3),
                        scalepoints  = TRUE,
                        title        = NULL)

shapeplot$plot <- shapeplot$plot + theme(plot.margin = margin(30, 30, 30, 30, "mm"))

resultsA <- dplyr::filter(ckbplotr_forest_data, name == "A")
forestplot <- make_forest_plot(panels            = list(resultsA),
                               col.key           = "variable",
                               row.labels        = ckbplotr_row_labels,
                               row.labels.levels = c("heading", "subheading", "label"),
                               rows              = c("Triglycerides concentration",
                                                     "Lipoprotein particle concentration"),
                               exponentiate      = TRUE,
                               panel.headings    = c(""),
                               xlim              = c(0.9, 1.1),
                               xticks            = c(0.9, 1, 1.1),
                               scalepoints       = TRUE,
                               pointsize         = 3,
                               plot.margin       = margin(8, 16, 30, 16, "mm"))


# Arrange the plots side-by-side
plots <- gridExtra::arrangeGrob(shapeplot$plot, forestplot$plot, nrow = 1)


# Create a text graphical object for the title (place it one line away from top left corner)
title <- grid::textGrob("Figure: Example of placing two plots on a page",
                        gp = grid::gpar(cex = 1, fontface = "bold"),
                        x = unit(0, "npc") + unit(1, "lines"),
                        y = unit(1, "npc") - unit(1, "lines"),
                        hjust = 0,
                        vjust = 1)

## Create a text graphical object for the footer (place it two line away from bottom left corner)
footer <- grid::textGrob("Footer",
                         gp = grid::gpar(cex = 0.7), 
                         just = c(0, 0),
                         x = unit(2, "lines"),
                         y = unit(2, "lines"))

## Arrange the title, plots, and footer vertically
figure <- gridExtra::arrangeGrob(title, plots, footer,
                                 nrow = 3,
                                 heights = c(10, 86, 4))

## Save the figure to a PDF, using landscape A4 dimensions
ggsave("example_page_layout.pdf", plot = figure, width = 11.69, height = 8.27, title = "Figure")
```

