---
title: "Save plots to files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Save plots to files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

```{r setup, include = FALSE}
library(ckbplotr)
```

## ggsave
The easiest way to save a plot is to to use the `ggplot2::ggsave()` function. Remember that `shape_plot()` and `forest_plot()` both return a list of the plot and code, so you need to use `$plot` to save just the plot to a file.
```{r}
my_results <- data.frame(
  x   = c(  12,   14, 15.5,   18),
  est = c(0.05, 0.21, 0.15, 0.32),
  se  = c(0.05, 0.05, 0.05, 0.05)
)

my_plot <- shape_plot(my_results,
                      xlims = c(10, 20),
                      ylims = c(0.75, 2),
                      exponentiate = TRUE,
                      quiet = TRUE)
```

```{r, eval = FALSE}
ggsave("myplot.png",
       plot = my_plot$plot,
       width = 14, height = 14, units = "cm")
```

Plots created with this package have transparent backgrounds, so if you're saving to a png file and want a solid white background (or generally want to change the background colour of a plot) use `theme()`:
```{r, eval = FALSE}
my_plot_with_bg <- my_plot$plot + 
  theme(plot.background = element_rect(fill = "white", colour = NA))
ggsave("myplot.png",
       plot = my_plot_with_bg,
       width = 14, height = 14, units = "cm")
```

The ckbplotr function `ggpreview()` can be used in place of `ggsave()` to preview the output.

## Save with title and footnote
Use `save_figure()` to add a title and footer to a plot and save to a file. The following code will save the plot (sized to 14 by 14 cm) in an A4 sized PDF file, with title and footer.
```{r save-shape-plot, eval = FALSE}
save_figure(my_plot$plot,
            filename   = "Figure 1.pdf",
            title      = "Figure 1: My example shape plot",
            footer     = "An example footer text.",
            size       = unit(c(14, 14), "cm"))
```

```{r, fig.width=8.27, fig.height=11.69, out.width='40%', out.extra='style="margin: auto; display: block; box-shadow: rgba(100, 100, 111, 0.2) 0px 7px 29px 0px;"', echo = FALSE}
figure <- prepare_figure(my_plot$plot,
                         title  = "Figure 1: My example shape plot",
                         footer = "An example footer text.",
                         size   = unit(c(14, 14), "cm"))
grid::grid.draw(figure$page)
```

The function has several arguments that allow for customization of appearance and layout. For example: `landscape = TRUE` to create a landscape page; and set `cropped` to be a file name to also save a plot without additional margins, title or footer.

## Preview the output
Set `preview = TRUE` to view a preview of the output in the RStudio Viewer pane. (Instead of creating a file, the figure is saved to a temporary PNG file and shown.)
```{r preview-shape-plot, eval = FALSE}
save_figure(my_plot$plot,
            filename   = "Figure 1.pdf",
            title      = "Figure 1: My example shape plot",
            footer     = "An example footer text.",
            size       = unit(c(14, 14), "cm"),
            preview    = TRUE)
```


