---
title: "Using special characters and formatting"
output:
  rmarkdown::html_vignette:
    fig_width: 7
    fig_height: 7
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{Using special characters and formatting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)

library(ckbplotr)

set.seed(57911624)
exampleresults <- function(){
  data.frame(variable = c('nmr_l_vldl_p', 'nmr_m_vldl_p', 'nmr_s_vldl_p',
                          'nmr_idl_p',
                          'nmr_l_ldl_p', 'nmr_m_ldl_p', 'nmr_s_ldl_p',
                          'nmr_l_vldl_tg', 'nmr_m_vldl_tg', 'nmr_s_vldl_tg',
                          'nmr_idl_tg',
                          'nmr_l_ldl_tg', 'nmr_m_ldl_tg', 'nmr_s_ldl_tg',
                          'nmr_l_vldl_c', 'nmr_m_vldl_c', 'nmr_s_vldl_c',
                          'nmr_idl_c',
                          'nmr_l_ldl_c', 'nmr_m_ldl_c', 'nmr_s_ldl_c'),
             estimate = rnorm(21, 0, 0.02),
             stderr   = 0.012 + abs(rnorm(21, 0, 0.015)),
             n        = round(runif(21, 100, 2000)),
             nb       = round(runif(21, 100, 2000)),
             shape    = rep(15, 21),
             colour   = "black",
             fill     = "black",
             bold     = FALSE,
             diamond  = FALSE,
             ciunder  = FALSE,
             stringsAsFactors = FALSE)
}
resultsA <- exampleresults()
resultsB <- exampleresults()

varheadings <- data.frame(heading1 = rep(c("Lipoprotein particle concentration",
                                           "Triglycerides concentration",
                                           "Cholesterol concentration"), each = 7),
                          heading2 = rep(c("VLDL", "VLDL", "VLDL",
                                           "IDL",
                                           "LDL", "LDL", "LDL"), times = 3),
                          heading3 = rep(c("Large", "Medium", "Small",
                                           NA,
                                           "Large", "Medium", "Small") , times = 3),
                          variable = c('nmr_l_vldl_p', 'nmr_m_vldl_p', 'nmr_s_vldl_p',
                                       'nmr_idl_p',
                                       'nmr_l_ldl_p', 'nmr_m_ldl_p', 'nmr_s_ldl_p',
                                       'nmr_l_vldl_tg', 'nmr_m_vldl_tg', 'nmr_s_vldl_tg',
                                       'nmr_idl_tg',
                                       'nmr_l_ldl_tg', 'nmr_m_ldl_tg', 'nmr_s_ldl_tg',
                                       'nmr_l_vldl_c', 'nmr_m_vldl_c', 'nmr_s_vldl_c',
                                       'nmr_idl_c',
                                       'nmr_l_ldl_c', 'nmr_m_ldl_c', 'nmr_s_ldl_c'),
                          stringsAsFactors = FALSE)
```

It is not always possible to use special characters or apply formatting to headings and other text directly in the plots produced by `make_shape_plot` and `make_forest_plot`. However, there are ways to create plots with the text (and formatting) you need. The two main methods are described below followed by some examples.

## Copy and edit the ggplot2 code
Both `make_shape_plot` and `make_forest_plot` return a character vector of ggplot2 code that is used to produce the plot. The code is also displayed in the Viewer pane in RStudio. You can copy this code, edit as needed, then use it in your R scripts directly. See the examples below for hints on where and how you need to edit the code.

## Add layers to a plot
If you want more control over some text on a plot but to still use a `make_*` function, you can add layers to the plot afterwards. I suggest you run the function first with some placeholder text, note which `geom_text` (or `annotate`) code is responsible for the text, rerun without that text (or replace it with ""), then add your own `geom_text`.


## Examples

### Change headings in make_forest_plot using expression
If you want to use different formatting and characters for row headings in `make_forest_plot` you can you the 'copy and edit' method. Just replace the section of code that creates the headings vector. For example you could replace it by a vector of expressions.

First start by using `make_forest_plot`.
```{r}
forestplot <- make_forest_plot(headings         = varheadings,
                               rows             = c("Triglycerides concentration"),
                               cols             = list(resultsA, resultsB),
                               exponentiate     = TRUE,
                               colnames         = c("Analysis A", "Analysis B"),
                               col.key          = "variable",
                               ci.delim         = " - ",
                               xlim             = c(0.9, 1.1),
                               xticks           = c(0.9, 1, 1.1),
                               blankrows        = c(1, 1, 0, 1),
                               scalepoints      = TRUE,
                               pointsize        = 3,
                               col.left         = c("n"),
                               col.left.space   = c(0.02),
                               col.left.heading = c("No. of\nevents"),
                               col.right.space  = 0.02,
                               col.heading.space = 1.5,
                               heading.space    = 2,
                               plot.space       = 8,
                               printplot        = FALSE)
```

Now copy the code from the RStudio Viewer pane, and edit the section under the comment "# Get a character vector of the headings, so these can be used in the plot". (Also edit the third line so that `datatoplot` is `forestplot$data`.)
```{r}
# ggplot2 code ------------------

# Prepare data to be plotted using ckbplotr::make_forest_data()
datatoplot <- ckbplotr::make_forest_data(
                headings = varheadings,
                rows = "Triglycerides concentration",
                cols = list(resultsA, resultsB),
                colnames = c("Analysis A", "Analysis B"),
                col.key = "variable",
                col.estimate = "estimate",
                col.stderr = "stderr",
                col.lci = NULL,
                col.uci = NULL,
                col.left = "n",
                col.right = NULL,
                col.keep = NULL,
                ci.delim = " - ",
                exponentiate = TRUE,
                blankrows = c(1, 1, 0, 1),
                scalepoints = TRUE,
                minse = NULL,
                addtext = NULL)

# Get a character vector of the headings, so these can be used in the plot
headings <- expression(bolditalic("Triglycerides concentration"),
                       "",
                       "VLDL",
                       "Large",
                       "Medium",
                       "Small",
                       "",
                       bold("IDL"),
                       "",
                       "LDL",
                       italic("Large"),
                       italic("Medium"),
                       italic("Small"),
                       "",
                       "")

# Get a character vector of the style for headings
boldheadings <- datatoplot %>%
                  dplyr::group_by(row) %>%
                  dplyr::summarise(bold = dplyr::if_else(all(is.na(estimate_transformed) | all(key %in% NULL)), "bold", "plain")) %>%
                  dplyr::arrange(row) %>%
                  dplyr::pull(bold)

# Identify any CIs that extend outside axis limits
datatoplot <- datatoplot %>%
                dplyr::mutate(cioverright  = (uci_transformed > 1.1),
                              uci_transformed = pmin(uci_transformed, 1.1),
                              cioverleft  = (lci_transformed < 0.9),
                              lci_transformed = pmax(lci_transformed, 0.9))

# Create the ggplot
ggplot(datatoplot, aes(x=-row, y=estimate_transformed)) +

  # Put the different columns in side-by-side plots using facets
  facet_wrap(~column, nrow = 1) +

  # Add a line at null effect
  annotate(geom = "segment", x=-1, xend=-Inf, y=1, yend=1, size = 0.5) +

  # Plot points at the transformed estimates
  ## Scale by inverse of the SE
  geom_point(aes(size = size, shape = 15, colour = "black", fill = "black"),
             stroke = 0.5,
             na.rm = TRUE) +

  # Scale the size of points by their side length
  # and make the scale range from zero upwards
  scale_radius(limits = c(0, 1),
               range = c(0, 3)) +

  # Plot the CIs
  geom_linerange(data = ~ dplyr::filter(.x, !is.na(estimate_transformed)),
                 aes(ymin = lci_transformed,
                     ymax = uci_transformed,
                     colour = "black"), 
                 size = 0.5,
                 na.rm = TRUE) +

  # Add tiny segments with arrows when the CIs go outside axis limits
  geom_segment(data = ~ dplyr::filter(.x, cioverright == TRUE),
               aes(x=-row,
                   y=uci_transformed-0.000001,
                   xend=-row,
                   yend=uci_transformed),
               colour = "black",
               size = 0.5,
               arrow = arrow(type = "closed", length = unit(4, "pt")),
               na.rm = TRUE) +
  geom_segment(data = ~ dplyr::filter(.x, cioverleft == TRUE),
               aes(x=-row,
                   y=lci_transformed+0.000001,
                   xend=-row,
                   yend=lci_transformed),
               colour = "black",
               size = 0.5,
               arrow = arrow(type = "closed", length = unit(4, "pt")),
               na.rm = TRUE) +

  # Use identity for aesthetic scales
  scale_shape_identity() +
  scale_fill_identity() +
  scale_colour_identity() +

  # Flip x and y coordinates
  coord_flip(clip = "off",
             ylim = c(0.9, 1.1)) +

  # Add columns to right side of plots
  ## column textresult
  geom_text(aes(x = -row, y = 1.104423626284,
            label = textresult),
            hjust = 0,
            size = 3,
            na.rm = TRUE,
            parse = TRUE) +
  annotate(geom = "text",
           x = 1.5, y = 1.104423626284,
           label = "HR (95% CI)",
           hjust = 0,
           size = 3,
           fontface = "bold") +

  # Add columns to left side of plots
  ## column n
  geom_text(aes(x = -row, y = 0.896395166165542,
                label = `n`,
            fontface = "plain"),
           hjust = 1,
           size = 3,
            na.rm = TRUE) +
  annotate(geom = "text",
           x = 1.5, y = 0.896395166165542,
           label = "No. of
events",
           hjust = 1,
           size = 3,
           fontface = "bold") +

  # Add xlab below each axis
  geom_text(aes(x = -Inf, y = 0.99498743710662, label = xlab),
            hjust = 0.5,
            size  = 3,
            vjust = 4.4,
            fontface = "bold",
            data = dplyr::tibble(column = factor(c("Analysis A", "Analysis B"),
                                                 levels = c("Analysis A", "Analysis B"),
                                                 ordered = TRUE),
                                 xlab = "HR (95% CI)")) +

  # Add column name above each column
  geom_text(aes(x = 1.5, y = 0.99498743710662, label = title),
            hjust = 0.5,
            nudge_x = 2,
            size  = 3,
            fontface = "bold",
            data = dplyr::tibble(column = factor(c("Analysis A", "Analysis B"),
                                                 levels = c("Analysis A", "Analysis B"),
                                                 ordered = TRUE),
                                 title = c("Analysis A", "Analysis B"))) +

  # Set the scale for the y axis (the estimates and CIs)
  scale_y_continuous(trans  = "log",
                     breaks = c(0.9, 1, 1.1),
                     expand = c(0,0)) +

  # Set the scale for the x axis (the rows)
  scale_x_continuous(breaks = -1:-max(datatoplot$row),
                     labels = headings,
                     name   = "",
                     expand = c(0,0)) +
  # Add the title
  labs(title = "") +

  # Control the overall looks of the plots
  theme(text             = element_text(size = 11),
        line             = element_line(size = 0.5),
        panel.background = element_rect(fill = "white", colour = NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line.x      = element_line(size = 0.5, lineend = "round"),
        axis.title.x     = element_blank(),
        axis.ticks.x     = element_line(colour = "black"),
        axis.text.x      = element_text(colour = "black",
                                        margin = margin(t = 4.4),
                                        vjust  = 1),
        axis.ticks.y     = element_blank(),
        axis.line.y      = element_blank(),
        axis.text.y      = element_text(hjust  = 0,
                                        colour = "black",
                                        face   = boldheadings,
                                        margin = margin(r = 2, unit = "lines")),
        panel.border     = element_blank(),
        panel.spacing    = unit(8, "lines"),
        strip.background = element_blank(),
        strip.placement  = "outside",
        strip.text       = element_blank(),
        legend.position  = "none",
        plot.margin      = unit(c(2,6,2,0), "lines"))
```

### Special characters
You can include some special characters just by putting their unicode in your text. But sometimes this does not work correctly, for example on Windows the code for a 'greater than or equal to' symbol "\\u2265" gets replaced by an equals sign. This can be fixed by using it in your own `geom_text` code (use either the 'copy and edit' or 'add layers' method).

Use a unicode symbol in column headings:
```{r}
forestplot <- make_forest_plot(headings         = varheadings,
                               rows             = c("Triglycerides concentration"),
                               cols             = list(resultsA, resultsB),
                               exponentiate     = TRUE,
                               colnames         = c("A", "B"),
                               colheadings      = c("", ""),
                               col.key          = "variable",
                               ci.delim         = " - ",
                               xlim             = c(0.9, 1.1),
                               xticks           = c(0.9, 1, 1.1),
                               blankrows        = c(1, 1, 0, 1),
                               scalepoints      = TRUE,
                               pointsize        = 3,
                               col.left         = c("n"),
                               col.left.space   = c(0.02),
                               col.left.heading = c("No. of\nevents"),
                               col.right.space  = 0.02,
                               col.heading.space = 1.5,
                               heading.space    = 2,
                               plot.space       = 8,
                               printplot        = FALSE)

forestplot$plot +
  geom_text(aes(x = 1.5, y = 0.99498743710662, label = title),
            hjust = 0.5,
            nudge_x = 2,
            size  = 3,
            fontface = "bold",
            data = dplyr::tibble(column = factor(c("A", "B"),
                                                 levels = c("A", "B"),
                                                 ordered = TRUE),
                                 title = c("A \u2265 10", " A < 10")))
```

### Formatting
Various formatting (called "aesthetics") such as colour, bold, italics, and angle can be controlled in `geom_text`. Make use of the arguments of `geom_text` (https://ggplot2.tidyverse.org/reference/geom_text.html) as you need.

Make column headings bold and italic:
```{r}
forestplot$plot +
  geom_text(aes(x = 1.5, y = 0.99498743710662, label = title),
            hjust = 0.5,
            nudge_x = 2,
            size  = 3,
            fontface = "bold.italic",
            data = dplyr::tibble(column = factor(c("A", "B"),
                                                 levels = c("A", "B"),
                                                 ordered = TRUE),
                                 title = c("A \u2265 10", " A < 10")))
```

### Expressions
The `geom_text` function has an argument `parse` which tells ggplot2 to parse text into expressions and display as described in [plotmath](https://rdrr.io/r/grDevices/plotmath.html). (Note that text in columns given by `col.right` in `make_forest_plot` already use `parse = TRUE`.)

Evaluate column headings as expressions:
```{r}
forestplot$plot +
  geom_text(aes(x = 1.5, y = 0.99498743710662, label = title),
            hjust = 0.5,
            nudge_x = 2,
            size  = 3,
            parse = TRUE,
            data = dplyr::tibble(column = factor(c("A", "B"),
                                                 levels = c("A", "B"),
                                                 ordered = TRUE),
                                 title = paste0("bold(", c("A>=10", "A^b"), ")")))
```
Note the expressions will remove the `fontface` and other formatting, but we can use `bold` inside the text instead (though it does not apply to symbols or numbers).


### ggtext
The [ggtext packge](https://wilkelab.org/ggtext/) adds support for some Markdown and HTML. You could install this package and use `geom_richtext()` in place of `geom_text()` to use some markdown (or HTML).

Bold and italics using markdown:
```{r}
forestplot$plot +
  ggtext::geom_richtext(aes(x = 1.5, y = 0.99498743710662, label = title),
                        hjust = 0.5,
                        size  = 3,
                        nudge_x = 2,
                        data = dplyr::tibble(column = factor(c("A", "B"),
                                                             levels = c("A", "B"),
                                                             ordered = TRUE),
                                             title = c("*A*", "**B<sup>2</sup>**")),
                        fill = NA,
                        label.color = NA)
```

You could also use `element_markdown` inside `theme` to make use of markdown just in row headings.
```{r, eval = F}
varheadings <- data.frame(heading1 = rep("Triglycerides concentration", each = 7),
                          heading2 = c("VLDL^a", "VLDL^a", "VLDL^a",
                                       "**IDL**",
                                       "*LDL*", "*LDL*", "*LDL*"),
                          heading3 = c("Large", "Medium", "Small",
                                       NA,
                                       "*Large*", "*Medium*", "*Small*"),
                          variable = c('nmr_l_vldl_tg', 'nmr_m_vldl_tg', 'nmr_s_vldl_tg',
                                       'nmr_idl_tg',
                                       'nmr_l_ldl_tg', 'nmr_m_ldl_tg', 'nmr_s_ldl_tg'),
                          stringsAsFactors = FALSE)

forestplot <- make_forest_plot(headings         = varheadings,
                               rows             = c("Triglycerides concentration"),
                               cols             = list(resultsA, resultsB),
                               exponentiate     = TRUE,
                               colnames         = c("A", "B"),
                               colheadings      = c("Analysis A", "Analysis B"),
                               col.key          = "variable",
                               ci.delim         = " - ",
                               xlim             = c(0.9, 1.1),
                               xticks           = c(0.9, 1, 1.1),
                               blankrows        = c(1, 1, 0, 1),
                               scalepoints      = TRUE,
                               pointsize        = 3,
                               col.left         = c("n"),
                               col.left.space   = c(0.02),
                               col.left.heading = c("No. of\nevents"),
                               col.right.space  = 0.02,
                               col.heading.space = 1.5,
                               heading.space    = 2,
                               plot.space       = 8,
                               printplot        = FALSE,
                               showcode         = FALSE)

forestplot$plot + 
  theme(axis.text.y = ggtext::element_markdown())
```



